---
title: Get Out The Vote
author: "Jim Gruman"
date: '`r format(Sys.Date())`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(tidyverse)
library(skimr)
library(viridis)
library(broom)
library(tidymodels)
library(janitor)
theme_set(theme_light())
```

## Get Out The Vote Case Study

Adapted from the Julia Silge [Supervised ML Course](https://supervised-ml-course.netlify.com/chapter3) and updated with Tidymodels in place of Caret.

Objective: use data on attitudes and beliefs in the United States to predict the 2016 voter turnout. Apply skills in dealing with imbalanced data and explore resampling options.

The Data Source is the [Democracy Fund Voter Survey ](https://www.voterstudygroup.org/publication/2016-voter-survey)

![Democracy Fund](https://www.voterstudygroup.org/assets/build/img/main-logo.svg)

```{r Download, message=FALSE, warning=FALSE, include=FALSE}

url<-"https://www.voterstudygroup.org/download?lv=xQ%252Fn%252FcP9yYLkNIyfm3hTNqbuYR%252BbJKplmR4kObnW5b9Qeh7rVivbHwCZLkuk2ceXKUCjqRTLwdFNwIYLAH7uYfTE%252B1dH7Xw%252FZh4Va%252Bme3J%252B7ak7fwiYryH7fWfSOwPmBSMxoLD9W02xhJMNfYMaSMwB9xw%252BIP63HhG6UzDxwLpazUHCuruvUykpJP3D2x0Jig9LbKMZ50h0SvWNGf3KEAg6YqLZP%252Fnvlg48HiTJRKCggIcxCYz%252B6udwwev4vaZ4MxyfgZZP79MoHb4T4lYY7r1mJMCSYYJhyTSqjieXnPsyGe16YRn5edhihsUEPl4yA"

download.file(url = url,
              destfile = "./dfvsg_2016_VOTER_Survey-csv.zip",
              mode = "wb")

unzip("dfvsg_2016_VOTER_Survey-csv.zip", files = "VOTER_Survey_December16_Release1.csv")

cols<-c(12,62:79,95:180)

voters<-read_csv("./VOTER_Survey_December16_Release1.csv", guess_max = 4000) %>% clean_names()

voters<-voters[,cols]

file.remove(c("./VOTER_Survey_December16_Release1.csv","./dfvsg_2016_VOTER_Survey-csv.zip" ))

voters<-voters %>%
  filter(!is.na(turnout16_2016)) %>%
  mutate(turnout16_2016 = fct_recode(as_factor(turnout16_2016),
                                       "Did.not.vote"="2",
                                       "Voted"="1"))
```
```{r}
voters %>%
  count(turnout16_2016)
```

Views captured in the survey included: 

1. Life in America today for people like you compared to fifty years ago is better? about the same? worse?

1. Was your vote primarily a vote in favor of your choice or was it mostly a vote against his/her opponent?

1. How important are the following issues to you?

- Crime
- Immigration
- The environment
- Gay rights

How do the reponses on the survey vary with voting behavior, at least on three survey metrics?

```{r warning=FALSE}

voters %>%
    group_by(turnout16_2016) %>%
    summarise(`Elections don't matter` = mean(rigged_system_1_2016 <= 2, na.rm=TRUE),
              `Economy is getting better` = mean(econtrend_2016 == 1, na.rm = TRUE),
              `Crime is very important` = mean(imiss_a_2016 == 2, na.rm = TRUE))

```


```{r warning=FALSE}
library(ggplot2)
library(ggtext)
scales::show_col(viridis_pal()(2))

voters %>%
  ggplot(aes(econtrend_2016, ..density.., fill = turnout16_2016))+
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) +
  labs(title = "Overall, is the economy getting *better* or *worse* for those that <b><span style='color:#440154FF'>Voted</span></b> and those that <b><span style='color:#FDE725FF'>Did not vote</span></b>?", 
       caption = str_c("Jim Gruman, ", Sys.Date()),
       x = element_blank()) +
  theme(legend.position = "none"  ,
        plot.title.position = "plot",
         plot.title = element_textbox_simple(
      size = 10,
      padding = unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
      margin = unit(c(0, 0, 5.5, 0), "pt"),
      fill = "white"
    ))+     #tip from William R Chase at RStudio Conference 2020
  scale_fill_viridis(discrete = TRUE)

```

Check the data missing-ness and find features with nearly zero variance

```{r message=FALSE, warning=FALSE}
Amelia::missmap(voters[,-1])

#library(GGally)
#ggscatmat(voters, columns = 1:104, color = "asia", alpha = 0.5)
```

Split the data into training and testing, then center, scale, and impute the missing fields with the recipes package functions. All the predictors in this dataset are numeric, so no dummies are required. If dummies were required, they would be created ahead of the centering, scaling, and box-cox transformations.

```{r}
voters_split<-voters %>% initial_split(strata = turnout16_2016)
voters_tr<-training(voters_split)
voters_te<-testing(voters_split)

voters_rec <- recipe(turnout16_2016 ~ .,
                    data = voters_tr) %>%
              step_knnimpute(all_predictors(), neighbors = 3) %>%
              step_center(all_predictors()) %>% 
              step_scale(all_predictors()) %>%
              check_missing(all_predictors()) %>%
              prep() 

# the steps taken
voters_rec

# what the data looks like after pre-processing
juice(voters_rec) 

x_train<- voters_rec %>% juice(all_predictors())
y_train<- voters_rec %>% juice(all_outcomes())

test_data<- bake(voters_rec, new_data=voters_te)

```

First fit a simple model, to take a quick look under the hood

```{r}

glm_spec<-logistic_reg(mode = "classification") %>%
         set_engine("glm")

glm_fit<- glm_spec %>%
          fit(turnout16_2016 ~ ., data = train_data)

rf_spec<-rand_forest(mode = "classification") %>%
         set_engine("ranger")

rf_fit<- rf_spec %>%
          fit(turnout16_2016 ~ ., data = train_data)

results_train <- glm_fit %>%
   predict(new_data = train_data) %>%
   mutate(
     truth = train_data$turnout16_2016,
     model = "glm"
   ) %>%
   bind_rows(rf_fit %>%
            predict(new_data = train_data) %>%
            mutate(truth = train_data$turnout16_2016,
                   model = "rf"
  )) 

results_test <- glm_fit %>%
   predict(new_data = test_data) %>%
   mutate(
     truth = test_data$turnout16_2016,
     model = "glm"
   ) %>%
   bind_rows(rf_fit %>%
            predict(new_data = test_data) %>%
            mutate(truth = test_data$turnout16_2016,
                   model = "rf"
  )) 

results_train %>%
  group_by(model) %>%
  metrics(truth = truth, .pred_class)

results_test %>%
  group_by(model)  %>%
  metrics(truth = truth, .pred_class)

```



Let's look for improvement with hyperparameter tuning on the random forest models.

```{r}
vote_boot <- bake(voters_rec, voters) %>%
        bootstraps(times = 10)

rf_spec <- rand_forest(
  mode = "classification",
  mtry = tune(),
  trees = 500,
  min_n = tune()
) %>%
  set_engine("ranger")

rf_grid <- tune_grid(
  rf_spec, 
  turnout16_2016 ~ .,
  resamples = vote_boot)

rf_grid %>%
  collect_metrics()

rf_grid %>%
  show_best("roc_auc")

```

So, choosing the random forest model with parameters `mtry = 5` and `min_n = 18` and assessing performance on testing data:

```{r}
rf_spec<-rand_forest(mode = "classification",
                     mtry = 5,
                     trees = 500,
                     min_n = 18) %>%
         set_engine("ranger")

rf_fit<- rf_spec %>%
          fit(turnout16_2016 ~ ., data = train_data)

rf_fit %>% predict(new_data = train_data) %>%
            mutate(truth = train_data$turnout16_2016,
                   model = "rf")%>%
  group_by(model)  %>%
  metrics(truth = truth, .pred_class)

```




