---
title: "WritingProjects"
author: "Jim Gruman"
date: '`r format(Sys.Date())`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Any new VM -----

R setup for any new Virtual Machine


```{r}
## at a command line, as sudo
## Install the  https://r-spatial.github.io/sf/  libudunits2-dev libgdal-dev libgeos-dev libproj-dev 

library(tidyverse)

library(tidymodels)

library(tinytex)
tinytex::install_tinytex()

library(blogdown)
blogdown::install_hugo()

library(sf)
library(mapview)
library(leaflet)
```

# Topics that I'd like to write articles about ----

Learning Resources for more advanced R 
Learning Resources for machine learning
Platform Learning (Latest from Azure)

Favorite Blogs and Podcasts



# TidyTuesday opportunities -----

AmesHousing with GLMNet Lasso, from the Chicago R Users Group

Geospatial -- USDA
Geospatial -- weather
Geospatial -- observations / descriptions
Geospatial -- time series / predictive /     often to display an area of predicted interest (due to temp/climate/water factors)

Football stadium attendance regression

GetOutTheVote did this person vote?






