---
title: "Brain Injury"
author: "Jim Gruman"
date: '`r format(Sys.Date())`'
output: github_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidymodels)
library(tidyverse)
library(vip)
theme_set(theme_minimal())
```

## R Markdown

```{r}
# Get the Data

tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
tbi_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_year.csv')
tbi_military <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_military.csv')


```

# Data Exploration

Brain Injury Awareness Month, observed each March, was established 3 decades ago to educate the public about the incidence of brain injury and the needs of persons with brain injuries and their families (1). Caused by a bump, blow, or jolt to the head, or penetrating head injury, a traumatic brain injury (TBI) can lead to short- or long-term changes affecting thinking, sensation, language, or emotion.

The goal of this week's #TidyTuesday is to spread awareness for just how common TBIs are - both in civilian and military populations.

> One of every 60 people in the U.S. lives with a TBI related disability. Moderate and severe traumatic brain injury (TBI) can lead to a lifetime of physical, cognitive, emotional, and behavioral changes.


```{r}
tbi_year %>%
  ggplot(aes(factor(year), number_est, fill = factor(injury_mechanism)))+
  geom_col()+
  labs(title = "Brain Injury Incidence",
       subtitle = "US CDC",
       caption = paste0("Jim Gruman ", Sys.Date()))+
  theme(plot.title.position = "plot") +
  scale_y_continuous(labels = scales::comma)+
  xlab("Year")+ylab("Number")
```


```{r}
library(tidyverse)
library(gganimate)
library(gifski)
library(png)
library(viridis)

p <- tbi_year %>% 
  filter(injury_mechanism != "Total") %>% 
  ggplot(aes(x = injury_mechanism,y = number_est, fill = type )) +
  geom_bar(position = "dodge", stat = "identity") + 
  viridis::scale_fill_viridis(discrete = TRUE, option = "A" ) +  
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 20)) +
  coord_flip() +
  transition_states(year, transition_length = 10, state_length = 1) +
    labs(title = "Traumatic Brain Injury (TBI) Incidence in {closest_state}",
       subtitle = "US CDC", x = "Cause of Injury", y = "Count",
       caption = paste0("Jim Gruman ", Sys.Date()))+
  theme(plot.title.position = "plot") +  
  scale_y_log10(labels = scales::comma) 


animate(p,  width = 900, height = 750, end_pause = 50, renderer = gifski_renderer("gganimqdodge1.gif"))
```


