---
title: "NFL attendance"
author: "Jim Gruman"
date: '`r Sys.Date()`'
output: github_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```
Let's build a very simple model for [NFL attendance from the #tidytuesday data of 4 February 2020](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-04/readme.md)

## Explore data

Load the attendance and team standings data from Github

```{r }
attendance <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/attendance.csv')

standings <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/standings.csv')

attendance_joined<-attendance %>%
  left_join(standings, by = c("year", "team_name", "team"))

```

Explore the files and look for trends. In this boxplot visual, some teams certainly expect higher weekly attendance.

```{r}
attendance_joined %>%
  filter(!is.na(weekly_attendance))%>%
 ggplot(aes(fct_reorder(team_name, weekly_attendance), 
            weekly_attendance, 
            fill = playoffs))+
 geom_boxplot(outlier.alpha = 0.5)+
 scale_y_continuous(labels = scales::label_comma())+
  labs(title = "NFL Weekly Attendance",
      caption = paste0("Jim Gruman ",Sys.Date()),
      x = "", y = "Weekly Attendance")+
  theme(plot.title.position = "plot",
        legend.position = "bottom")+
    coord_flip()+
   theme_minimal()

```

In this histogram, playoff-bound teams generally have higher point spread margins over the course of many games.

```{r}
attendance_joined %>%
  distinct(team_name, year, margin_of_victory, playoffs)%>%
  ggplot(aes(margin_of_victory, fill = playoffs))+
  geom_histogram(position = "identity", alpha = 0.7)+
  labs(title = "NFL Team Margin of Victory",
      caption = paste0("Jim Gruman ",Sys.Date()),
      x = "Point Spread", y = "Game Count")+
  theme(plot.title.position = "plot",
        legend.position = "bottom")+
   theme_minimal()
```

Across the weeks of the season, this data visualization shows the distribution of attendance by week number.

```{r}
attendance_joined %>%
  mutate(week = factor(week))%>%
  ggplot(aes(week, weekly_attendance, fill = "blue"))+
  geom_boxplot(show.legend = FALSE, outlier.alpha = 0.4)+
  labs(title = "NFL Weekly Game Attendance",
      caption = paste0("Jim Gruman ",Sys.Date()),
      x = "Week of Season", y = "Attendance")+
  theme(plot.title.position = "plot",
        legend.position = "bottom")+
  scale_y_continuous(labels = scales::label_comma())+
   theme_minimal()

```

To build models for illustratign the prediction of weekly attendance, we will select on the team_name, the year, the week of the game, and the margin of victory.

```{r}
attendance_df<-attendance_joined  %>%
  filter(!is.na(weekly_attendance))%>%
  select(weekly_attendance, team_name, year, week, 
         margin_of_victory, strength_of_schedule, playoffs)

```

## Train a Model

First, the data are split into training and testing sets at about 75/25, retaining similar playoff outcomes in both.

```{r}
library(tidymodels)

attendance_split<-attendance_df %>%
  initial_split(strata = playoffs)

nfl_train <-training(attendance_split)
nfl_test <- testing(attendance_split)


```

A simple linear model is specified and fit, and the estimates of coefficients are shown here:

```{r}
lm_spec<- linear_reg(mode = "regression")%>%
  set_engine(engine = "lm")

lm_fit<-lm_spec %>%
  fit(weekly_attendance ~ ., data = nfl_train)

tidy(lm_fit) %>% arrange(-estimate)

```

A comparable random forest regression is specified and fit here:

```{r}
rf_spec<- rand_forest(mode = "regression") %>%
  set_engine(engine = "ranger")

rf_fit<-rf_spec %>%
  fit(weekly_attendance ~ ., data = nfl_train)

tidy(lm_fit) %>% arrange(-estimate)

```

## Evaluate Models



```{r}
results_train<-lm_fit %>%
  predict(new_data = nfl_train) %>%
  mutate(truth = nfl_train$weekly_attendance,
         model = "lm")%>%
  bind_rows(rf_fit %>%
      predict(new_data = nfl_train) %>%
      mutate(truth = nfl_train$weekly_attendance,
         model = "rf"))

results_test<-lm_fit %>%
  predict(new_data = nfl_test) %>%
  mutate(truth = nfl_test$weekly_attendance,
         model = "lm")%>%
  bind_rows(rf_fit %>%
      predict(new_data = nfl_test) %>%
      mutate(truth = nfl_test$weekly_attendance,
         model = "rf"))

```
```{r}
results_train %>%
  group_by(model)%>%
  rmse(truth = truth, estimate = .pred)
```

```{r}
results_test %>%
  group_by(model)%>%
  rmse(truth = truth, estimate = .pred)
```

The random forest model here appears to overfit the training data set, with disappointing results on new data.

```{r}

results_test %>%
  mutate(train = "testing") %>%
  bind_rows(results_train %>%
              mutate(train = "training")) %>%
  ggplot(aes(truth, .pred, color = model))+
  geom_abline(lty = 2, color = "gray80", size = 1.5)+
  geom_point(alpha = 0.5)+
  facet_wrap(~train)+
  theme_minimal()

```
### Lets try again, with resampling on the training

```{r}
library(furrr)

nfl_folds<- vfold_cv(nfl_train, strata = playoffs)

rf_res<-fit_resamples(
  weekly_attendance ~ .,
  rf_spec,
  nfl_folds,
  control = control_resamples(save_pred = TRUE)
)

rf_res %>% 
  collect_metrics()
```

```{r}
rf_res %>%
  unnest(.predictions) %>%
  ggplot(aes(weekly_attendance, .pred, color = id))+
  geom_abline(lty = 2, color = "gray80", size = 1.5)+
  geom_point(alpha = 0.5)+
  labs(title = "Model Accuracy",
      caption = paste0("Jim Gruman ",Sys.Date()),
      x = "Attendance Truth", y = "Attendance Prediction")+
  theme(plot.title.position = "plot",
        legend.position = "bottom")+
  scale_y_continuous(labels = scales::label_comma())+
  scale_x_continuous(labels = scales::label_comma())+
     theme_minimal()

```

Credits:
  Julia Silge, RStudio
  Thomas Mock, RStudio



