---
title: "BeerProduction"
author: "Jim Gruman"
date: "`Sys.Date`"
output: github_document
name: beer_production
owner: jim_gruman
metrics:
  total_pounds:
    title: "# of pounds in production"
    description: "Number of pounds used in Beer Production in the US"
dimensions:
  material_type:
    title: Type
    description: Grain or nonGrain Products
  material:
    title: Material
    description: Wheat, hops, malt, barley, etc.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)   
library(lubridate)
library(tidymetrics)
library(viridis)  
library(GGally)      
library(gghighlight) 
library(Cairo)      
library(tidymodels)
theme_set(theme_minimal())

```

## Import this week's data
```{r message=FALSE}
brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv') %>%
    mutate(date = ymd(paste0(year, "/", month, "/1"))) 
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
```

## Exploration

```{r eval = FALSE}
skimr::skim(beer_states)
# 19 year/states with missing barrels

skimr::skim(brewing_materials)
table(brewing_materials$type)
# opportunity to examine the proportional trends of Corn vresus Sugar/syrups use

skimr::skim(beer_taxed)

skimr::skim(brewer_size)
table(brewer_size$brewer_size)
```

Total all barrels produced by state per year (regardless of type)

```{r}
all_barrels <-
  beer_states %>%
  na.omit() %>%
  filter(state != 'total') %>%
  group_by(state, year) %>%
  summarise(barrels = sum(barrels))
```

The top 5 states with largest increase in barrels from 2008 to 2019

```{r}
increase_states <-
  all_barrels %>%
  mutate(total_barrels = cumsum(barrels)) %>%
  filter(year %in% c(2008, 2019)) %>%
  mutate(pct_chg = (total_barrels - lag(total_barrels)) / lag(total_barrels)) %>%
  ungroup() %>%
  top_n(5, pct_chg) %>%
  pull(state)     # Alabama, Connecticut, North Dakota, Rhode Island, Tennesee
```

```{r}
#Lets get the state names
states <- tibble(state = state.abb, state_name = state.name)

beer_df <- 
  beer_states %>%
  na.omit() %>%
  filter(state != 'total') %>%
  group_by(state, year) %>%
  summarise(barrels = sum(barrels)) %>%
  mutate(total_barrels = cumsum(barrels)) %>%
  left_join(states, by = 'state')
```

Lets have a look at the materials trends. What ingredients are used in beer production?

```{r}
brewing_materials %>%
  filter(!str_detect(material_type, "Total"),
         !(month == 12 & year %in% 2014:2015)) %>%  # remove the subtotals) %>%
  filter(date == max(date)) %>%
  mutate(type = fct_reorder(type, month_current)) %>%
  ggplot(aes(type, month_current/1000000, fill =  type)) +
  geom_col()+
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  coord_flip()+
  theme(plot.title.position = "plot",
        legend.position = "")+
  viridis::scale_color_viridis(discrete = TRUE)  +
  labs(
    x = 'Ingredient',
    y = 'Million Pounds',
    title = 'Ingredients in US Beer Production, in the month of December 2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )
```

```{r}
brewing_materials %>%
  filter(!str_detect(type, "Total"), year < 2016,
         !(month == 12 & year %in% 2014:2015)) %>%  # remove the subtotals
  ggplot(aes(date, month_current/1000000, color= type)) +
  geom_line()+
  viridis::scale_color_viridis(discrete = TRUE) +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )
```

Tidymetrics

```{r }
brewing_summarized <-brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total"), year < 2016) %>%
  cross_by_dimensions(material_type, material) %>%  # duplicates the data to summarize it
  cross_by_periods(c("month","quarter","year")) %>%
  summarize(total_pounds = sum(month_current)) %>%
  ungroup()

brewing_summarized %>%
  filter(material_type != "All", material != "All", period == "month") %>%
  ggplot(aes(date, total_pounds/1000000, color = material))+
  geom_line()+
  viridis::scale_color_viridis(discrete = TRUE) +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )

brewing_summarized %>%
  filter(material_type != "All", material != "All", period == "quarter") %>%
  ggplot(aes(date, total_pounds/1000000, color = material))+
  geom_line()+
  viridis::scale_color_viridis(discrete = TRUE) +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )

brewing_summarized %>%
  filter(material_type != "All", material != "All", period == "year") %>%
  ggplot(aes(date, total_pounds/1000000, color = material))+
  geom_line()+
  viridis::scale_color_viridis(discrete = TRUE) +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )  

```

```{r}
brewing_materials %>%
  filter(material_type == "Total Used", year < 2016) %>%
  mutate(month = factor(month, labels = month.abb)) %>%
  ggplot(aes(x= month,y=month_current/1000000, group = factor(year), color = factor(year)))+
  geom_line() +
  expand_limits(y=0) +
  viridis::scale_color_viridis(discrete = TRUE) +
    labs(x = 'Month',
    y = 'Million Pounds',
    color = "Year",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday') +  
  scale_y_continuous(labels = scales::comma_format())   

```

```{r eval=FALSE}
brewing_metrics<-create_metrics(brewing_summarized)

library(shinymetrics)

# interactive plotly visualization of grouped time multi-dimensional BI 
preview_metric(brewing_metrics$production_NA_total_pounds)
# change the dropdown from "last year" to "all time" because 2019 is not in this dataset
```

```{r}
brewer_size %>%
  filter(brewer_size != "Total", !is.na(total_barrels))%>%
  mutate(brewer_size = fct_lump(brewer_size, 5, w = total_barrels),
         barrel_number = coalesce(parse_number(as.character(brewer_size)),1),
         brewer_size = fct_reorder(brewer_size, barrel_number)) %>%
  ggplot(aes(as_factor(year), total_barrels, fill = brewer_size))+
  geom_col() +
  scale_y_continuous(labels = scales::comma_format())  +
  viridis::scale_fill_viridis(discrete = TRUE) +
    labs(x = 'Year',
    y = 'Total Barrels',
    fill = "Brewer Size",
    title = 'Total Production, by US Brewer Size in 2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday')   +
  theme(plot.title.position = "plot")
```

## Where is Beer produced?

```{r}
increase_plot <-
  ggplot(beer_df, aes(year, barrels, group = state)) +
  geom_line(aes(color = state), show.legend = F) +
  gghighlight(                               
    state %in% increase_states,
    label_key = state_name,
    label_params = list(size = 3, family = 'Arial Narrow',
                        fill = 'grey95', alpha = 0.8, 
                        label.r = 0.1, segment.alpha = 0),
    unhighlighted_params = list(color = 'grey88')
  ) +
  labs(
    x = 'Year',
    y = 'Total number of barrels produced',
    title = 'States with the Largest Increase in Beer Production, 2008-2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  ) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_log10(labels = scales::label_number()) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme(panel.grid = element_blank(),
        panel.grid.major = element_blank(),
        plot.title.position = "plot")

ggsave(increase_plot, file = 'Increase_StateBeerProd.png', type = 'cairo')
```

```{r}
beer_states %>%
  count(type, sort = TRUE, wt = barrels)

beer_states %>%
  group_by(year) %>%
  summarize(barrels = sum(barrels, na.rm = TRUE))

# Who consumes beer on premises?

beer_states %>%
  filter(type == "On Premises", year == max(year), state != "total")%>%
  arrange(desc(barrels))

beer_states %>%
  filter( year == max(year), state != "total")%>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "On Premises") %>%
  arrange(desc(percent))

beer_states %>%
  filter( year == max(year), state != "total")%>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "Bottles and Cans") %>%
  arrange(desc(percent))

beer_states %>%
  filter( year == max(year), state != "total")%>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "Kegs and Barrels") %>%
  arrange(desc(percent))

states_percents_2019 <-beer_states %>%
  filter( year == max(year), state != "total")%>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup() 

states_percents_2019 %>%
  filter(type == "On Premises") %>%
  arrange(desc(percent))

```

```{r}
library(maps)
library(sf)

states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))

states

 states_percents_2019 %>%
   mutate(ID = str_to_lower(state.name[match(state,state.abb)]))%>%
   inner_join(states, by = "ID") %>%
   ggplot()+
   scale_fill_gradient2(low = "yellow", high = "blue", midpoint = 0.5,
                        labels = scales::percent)+
   geom_sf(aes(geometry = geom, fill = percent)) +
   facet_wrap(~ type, nrow = 2)+
   ggthemes::theme_map()+
   labs(title = "In Each State, What is the Highest Proportion of Beer Consumed?",
        subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
        caption = '@Jim_Gruman | #TidyTuesday',
        fill = "% Percent")+
   theme(legend.position = c(0.7,0.1))
   
```

## An animamated map.  Because it's TidyTuesday

```{r}
library(gganimate)

p<-beer_states %>%
  filter(state != "total")%>%
  group_by(state, year) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup() %>%
  mutate(ID = str_to_lower(state.name[match(state, state.abb)]))%>%
  inner_join(states, by = "ID") %>%
  ggplot(aes(geometry = geom, fill = percent))+
  geom_sf() +
  transition_time(year) +
  facet_wrap(~ type, nrow = 2) +
  scale_fill_gradient2(low = "yellow", high = "blue", midpoint = 0.5,
                        labels = scales::percent) +
  ggthemes::theme_map() +
  labs(title = "In Each State, What is the Highest Proportion of Beer Consumed? {as.integer(frame_time)}",
        subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau, 2008-2019',
        caption = '@Jim_Gruman | #TidyTuesday',
        fill = "% Percent")+
   theme(legend.position = c(0.7,0.1))



animate(p,  width = 900, height = 750, end_pause = 50, renderer = gifski_renderer("gganimqdodge1.gif"))

```

## What is the relationship between the materials used? (using resampling methods)

# How much sugar do beer producers need per pound of malt?

```{r}
brewing_materials %>% count(type, wt = month_current,  sort = TRUE)

brewing_filtered<-brewing_materials %>%
  filter(!str_detect(type, "Total"), year < 2016,
         !(month == 12 & year %in% 2014:2015))

```

```{r}
brewing_df<-brewing_filtered %>%
  select(date, type, month_current) %>%
  pivot_wider(names_from = type, 
              values_from = month_current) %>%
  janitor::clean_names()

# is there a remote correlation between
brewing_df %>%
  ggplot(aes(malt_and_malt_products, sugar_and_syrups))+
  geom_smooth(method = "lm")+
  geom_point()
  
```


A simple linear model
```{r}
# force the y-intercept to be zero (no beer, no malt, no sugar)
beer_fit<-lm(sugar_and_syrups ~ 0 + malt_and_malt_products, data = brewing_df)

summary(beer_fit)
```

```{r}
tidy(beer_fit)
```

## Bootstrap Resampling

to arrive at a confidence interval for a linear model

```{r}
beer_boot<-bootstraps(data = brewing_df, times = 1e3, apparent = FALSE)
```

```{r}
beer_models<- beer_boot %>% 
  dplyr::mutate(model = purrr::map(splits, ~ lm(sugar_and_syrups ~ 0 + malt_and_malt_products, 
                                  data = . )),
  coef_info = purrr::map(model, tidy))

beer_coefs<-beer_models %>%
     unnest(coef_info)
     
```

## Evaluate results

What is the distribution of the estimates for the model?

```{r}
beer_coefs %>%
  ggplot(aes(estimate))+
  geom_histogram(alpha = 0.7)
```

Our bootstrap confidence intervals:

```{r}
int_pctl(beer_models, coef_info)
```


```{r}
beer_aug<-beer_models %>%
  dplyr::mutate(augmented = purrr::map(model, augment)) %>%
  unnest(augmented)

beer_aug %>%
  ggplot(aes(malt_and_malt_products, sugar_and_syrups))+
  geom_line(aes(y=.fitted, group = id), alpha = 0.1, color = "cyan3")+
  geom_point()+
  labs(title = "Bootstrap Resampling Model Estimates Confidence Interval")
```

