---
title: "BicyclePricing"
author: "Jim Gruman"
date: "3/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
theme_set(theme_minimal())
```

Insipired by Matt Dancho's [Lab 5: Hands-On Codign with the NEW parsnip package](https://university.business-science.io/courses/541207/lectures/9846389)

## Load bicycle pricing and feature data

As a first step, I scrape data from a neighborhood bike shop. This dataset includes a wide variety of bicycles, from kids sizes and BMX up through carbon fibre time trial and electric bikes.

```{r}
library(rvest)
library(janitor)
library(tidyverse)
# Modeling packages
library(tidymodels)
#library(parsnip)
library(rsample)
library(yardstick)
library(broom)
# Connector packages
library(rpart)
library(rpart.plot)
library(xgboost)

top_url <- "https://www.bicyclesetc-il.com"
pl_url<- "/product-list/bikes-1000/"

# build a product list
startrow <- as.character(seq(from = 0, to = 480, by =60))
fun<-function(x)(paste0(top_url, pl_url, "?maxItems=60&startrow=", x))
productlist_urls<- map(startrow, fun)

productlist<-map(productlist_urls, read_html) 

products<-productlist %>%
map(~html_nodes(., ".seItemName")) %>%
map(html_text) %>%
unlist()

pricelist<-productlist %>%
map(~html_nodes(., ".seRegularPrice, .seSpecialPrice")) %>%
map(html_text) %>%
map(~str_sub(., start = 2L)) %>%
map(~parse_number(.)) %>%
  unlist()

brandlist<-productlist %>%
map(~html_nodes(., ".seBrandName")) %>%
map(html_text) %>%
unlist()

pagelist<-productlist %>%
map(~html_nodes(., ".seProductAnchor")) %>%
map(~html_attr(., "href")) %>%
map(~paste0(top_url, .)) %>%
unlist()

#length(products)
#length(pricelist)
#length(brandlist)
#length(pagelist)

bikes<-tibble(product = products, price = pricelist, brand = brandlist)

bikes$ID <- seq.int(length(bikes$product))

# read detailed product specification pages

productdetail<-map(pagelist, read_html)

productdetail[[1]] %>%
    map(~html_nodes(., ".seProductBreadcrumb")) 


detail<-productdetail %>%
    map(~html_nodes(., "table.table.seProductSpecTable")) %>%
    map(~html_table(., fill = TRUE)) 

names(detail) <-seq.int(length(detail))

detail<-detail %>%
    map_dfr(1, .id = "ID")

detail$ID = as.integer(detail$ID)

bikes_df<-detail %>%
  left_join(bikes, by = "ID") %>% 
  select(product, price, brand, X1, X2) %>%
  pivot_wider(names_from = X1, values_from = X2, values_fn = list(X2 = max)) %>%
  clean_names() %>%
  mutate(rims_wheels = case_when(!is.na(rims_wheels) ~ rims_wheels,
                      !is.na(wheels_rims) ~ wheels_rims,
                      !is.na(rims) ~ rims,
                      TRUE ~ "unknown")) %>%
  select(-wheels_rims, -rims) %>%
  mutate(frame_spec = case_when(str_detect(tolower(frame), "carbon") ~ "carbon fiber",
       str_detect(tolower(frame), "smartform c1") ~ "smartform c1",
       str_detect(tolower(frame), "smartform c2") ~ "smartform c2",
       str_detect(tolower(frame), "smartform c3") ~ "smartform c3",
       str_detect(tolower(frame), "6061|6000|aluminum|superlight") ~ "aluminum",
       TRUE ~ "steel")) %>%
  mutate(brake_spec = case_when(str_detect(tolower(brakes), "coaster") ~ "coaster",
         str_detect(tolower(brakes), "caliper|linear|v-brake|u-brake") ~ "caliper",
         str_detect(tolower(brakes), "promax") ~ "Promax",
         str_detect(tolower(brakes), "mechanical disc") ~ "mechanical disc",
         str_detect(tolower(brakes), "rt900") ~ "RT900",
         str_detect(tolower(brakes), "rt800") ~ "RT800",
         str_detect(tolower(brakes), "sram level ultimate") ~ "SRAM Level Ultimate",
         str_detect(tolower(brakes), "sram force 1") ~ "SRAM Force 1",
         str_detect(tolower(brakes), "sram force etap") ~ "SRAM Force eTap",
         str_detect(tolower(brakes), "sram red etap") ~ "SRAM RED eTap",
         str_detect(tolower(brakes), "shimano xtr") ~ "shimano XTR",
         str_detect(tolower(brakes), "sram code rsc") ~ "SRAM Code RSC",
         str_detect(tolower(brakes), "sram level tlm") ~ "SRAM Level TLM",  
         str_detect(tolower(brakes), "sram apex") ~ "SRAM Apex",
         str_detect(tolower(brakes), "aero") ~ "Aero Integrated",
         str_detect(tolower(brakes), "mechanical disc|trp spyre") ~ "mechanical disc",
         str_detect(tolower(brakes), "hyd|shimano|hayes|tektro") ~ "hydraulic disc",
         TRUE ~ "other")) %>%
  select(product, price, brand, frame_spec, brake_spec, drive_system, display, motor)

```

## Exploration

A quick exploratory peek at the range of prices for each frame specification:

```{r}
ggplot(bikes_df, aes(frame_spec, price))+
  geom_boxplot()+
  labs(title = "Bicycle Pricing by Frame Specification",
       subtitle = "For a Neighborhood Bike Shop",
       caption = paste0("Jim Gruman ", Sys.Date()))+
  theme(plot.title.position = "plot",
        panel.grid.major.x = element_blank(),
        legend.position = "none") +
  scale_y_continuous(labels=scales::dollar_format())+
  xlab("")+ylab("Price")
```

## Modeling

The model we are going to build predicts price as a function of the features. The first step, as always, is to split the data.

```{r}
set.seed(42)
split_obj <- rsample::initial_split(bikes_df, 
                                    prop   = 0.8)

train_tbl <- split_obj %>% training() 
test_tbl  <- split_obj %>% testing() 
```

The `tidymodels` meta package provides a consistent access point to several machine learning algorithms 

# - Cross Validation & Grid Search:

# 4.1 Linear Regression (No Engineered Features) ----
`?linear_reg`     # Step 1: Pick a parsnip algorithm, & Set key parameters
`?set_engine`     # Step 2: Set an engine, Returns a model spec
`?fit.model_spec` # Step 3: Fit model specification to data




