---
title: "Tate Art Collection"
date: "January 12, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

The data this week comes from the [Tate Art Collections](https://github.com/tategallery/collection). It consists of metadata for around 70,000 artworks that Tate owns or jointly owns with the National Galleries of Scotland as part of [ARTIST ROOMS](http://www.tate.org.uk/artist-rooms). Metadata for around 3,500 associated artists is also included.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	cache.lazy = FALSE,
	df_print = "paged",
	dpi = 300,
	tidy = "styler"
)

library(tidyverse)
library(tidytuesdayR)
library(treemapify)
library(hrbrthemes)
library(countrycode)

extrafont::loadfonts(quiet = TRUE)

theme_set(hrbrthemes::theme_ft_rc() +
            theme(plot.title.position = "plot",
                  plot.caption.position = "plot"))


```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-01-12")

```

```{r Readme, eval = interactive()}

tt

```


```{r prepare}


art <- tt$artwork %>% 
  rename(art_url = url) %>% 
  left_join(tt$artists, by = c("artistId" = "id")) %>% 
  mutate(gender = factor(gender),
         on = str_sub(medium, start = str_locate(medium, " on ")[,2])) 

```


# Visualize

Treemap of Artists

```{r Visualize}

art %>% 
  mutate(birthDecade = yearOfBirth %/% 10 * 10) %>% 
  count(artist, birthDecade) %>% 
  slice_max(order_by = n, n = 60) %>% 
  filter(!is.na(birthDecade), artist != "Turner, Joseph Mallord William") %>% 
  ggplot(aes(area = n, fill = birthDecade, label = artist, subgroup = birthDecade)) +
  geom_treemap() +
  geom_treemap_subgroup_border() +
  geom_treemap_text(color = "black", place = "top", reflow = TRUE) +
  geom_treemap_subgroup_text(color = "gray40", place = "bottomright", 
                             fontface = "italic", min.size = 0, alpha = 0.3) +
  scale_fill_fermenter(palette = "YlOrRd", direction = -1,
                       guide = "colorsteps") +
  labs(fill = NULL, title = "The Tate Collection's Artists by birth year (Non-Turner)",
       subtitle = "Born in 1775, Joseph Mallord William Turner's works account for 57% alone. The size of\nthe box corresponds to the number of pieces of each non-Turner artist in the collection.",
       caption = "Data Source: Tate Art Collection")
  
```

And the Tidytuesday tweet:

```{r echo=FALSE}
blogdown::shortcode('tweet', '1349150540610076673')
```

