---
title: "US Broadband Access"
author: "Jim Gruman"
date: "May 5, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
    code_download: true
    df_print: paged
editor_options:
  chunk_output_type: console
---

The #TidyTuesday data this week is on Internet Access, from Microsoft by way of The Verge. 

> If broadband access was a problem before 2020, the pandemic turned it into a crisis. As everyday businesses moved online, city council meetings or court proceedings became near-inaccessible to anyone whose connection couldn’t support a Zoom call. Some school districts started providing Wi-Fi hotspots to students without a reliable home connection. In other districts, kids set up in McDonald’s parking lots just to get a reliable enough signal to do their homework. After years of slowly widening, the broadband gap became impossible to ignore.

First, let's load libraries and set a ggplot theme:

```{r setup, warning=FALSE}

options(tigris_use_cache = TRUE)
suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)
library(tidycensus)
  
library(sf)
library(patchwork)
})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))

```

Loading the data 

```{r load, message=FALSE}
broadband_il <- tidytuesdayR::tt_load("2021-05-11")$broadband %>%
  janitor::clean_names() %>%
  filter(st == "IL") %>% 
  transmute(GEOID = as.character(county_id),
            usage = parse_number(broadband_usage, na = "-" )) %>% 
  left_join(get_acs(geography = "county",
                    variables = "B19013_001",
                    state = "IL",
                    geometry = TRUE))
```

```{r}
plot_il <- function(variable) {
  broadband_il %>%
    st_as_sf() %>%
    ggplot(aes(fill = {{variable}} )) +
    geom_sf(color = "gray80", size = 0.2) +
    coord_sf(label_axes = "----",
             label_graticule = "----") +
    theme(panel.grid.major = element_blank())
}

plot_il(estimate) +
  scale_fill_viridis_b(option = "plasma",
                       n.breaks = 8,
                       labels = scales::comma) +
  labs(subtitle = "Median Household Income",
       fill = "$US") +
  plot_il(usage) +
  scale_fill_viridis_b(n.breaks = 8,
                       labels = scales::percent_format(accuracy = 1)) +
  labs(subtitle = "High Speed Broadband Usage",
       fill = NULL) +
  plot_annotation(
    title = "High Speed Broadband in Illinois",
    caption = "Data Sources: Microsoft and US Census ACS 2019 \n
                  Percent of people per county that use the internet at more than 25 Mbps/3 Mbps"
  ) &
  theme(plot.margin = margin (0.5, 0, 0, 0, "cm"),
        plot.title = element_text(size = 30))
```

Related tweets and inspirations:

```{r}
tweetrmd::include_tweet("https://twitter.com/jrosecalabrese/status/1392299000401858563")

tweetrmd::include_tweet("https://twitter.com/juliasilge/status/1392324410082689026")

tweetrmd::include_tweet("https://twitter.com/jim_gruman/status/1392486129786572801")

```


