---
title: "Video Games and Sliced"
date: "March 16, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
    code_download: true
    df_print: paged
editor_options:
  chunk_output_type: console
---

The #TidyTuesday data this week comes from Steam by way of Kaggle and originally came from SteamCharts. The data was scraped and uploaded to Kaggle.

```{r setup, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	cache.lazy = FALSE,
	df_print = "paged",
	dpi = 300,
	tidy = "styler",
	dev = "ragg_png",
	autodep = TRUE
)

suppressPackageStartupMessages({
library(tidyverse)
library(tidymodels)
library(lubridate)
library(hrbrthemes)
library(systemfonts)
 })

ggplot2::theme_set(hrbrthemes::theme_ft_rc() +
            ggplot2::theme(plot.title.position = "plot",
                           plot.caption.position = "plot"))

```

```{r}
tt <- tidytuesdayR::tt_load("2021-03-16")
```


```{r, eval = interactive()}
tt
```

```{r}
games <- tt$games %>% 
  mutate(date = ym(paste0(as.character(year), month))) %>% 
  complete(nesting(date, gamename), fill = list(avg = 0, peak = 0))
```

Let's have a look at the games themselves

## Common games {.tabset}

### Top Games

```{r}
games %>% 
  filter(avg > 0) %>% 
  count(gamename, sort = TRUE)
```

### Plot

```{r}
games %>% 
  filter(avg > 0) %>% 
  count(gamename, sort = TRUE) %>% 
  ggplot(aes(n)) +
  geom_histogram(bins = 30) +
  labs(y = "Count of Months in dataset",
       x = "Number of Titles")
```

## {-}

Discoveries:

1. There are some non-UTF characters in the game names.

1. There are 1248 distinct game names. 

1. About 200 of the games appear in all 104 months of the study(??)


```{r}
games %>%
  group_by(date) %>%
  summarise(engagement = sum(avg)) %>%
  ggplot(aes(date, engagement)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "year",
               date_labels = "%Y") +
  labs(x = NULL, y = NULL, title = "Steam Engagement, Sum of Monthly Avg", 
       caption = paste0("Source: Kaggle"))
```

The Sid Meier series:

```{r}
games %>%
  filter(str_detect(gamename, "Sid Meier")) %>%
  mutate(name = str_remove(gamename, "Sid Meier's Civilization"),
         name = str_remove(name, "\\: ")) %>% 
  ggplot(aes(date, peak, color = name)) +
  geom_line() +
  scale_color_brewer(name = NULL,
                    direction = -1,
                    palette = "BrBG") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "year",
               date_labels = "%Y") +
  labs(
    x = NULL,
    y = NULL,
    title = "Sid Meier's Civilization Peak Engagement on Steam",
    caption = paste0("Source: Kaggle")
  ) +
  theme(legend.position = c(0.8,0.75),
        legend.background = element_rect(color = "white"))
```


I wonder if the all of the game engagement has a seasonal peak, like at a given month?

```{r}
games %>%
  mutate(month = factor(month, levels = month.name),
         month = fct_relabel(month, ~ month.abb)) %>%
  group_by(year, month) %>%
  summarise(avg = median(avg),
            .groups = "drop") %>%
  filter(row_number() != first(row_number())) %>%
  mutate(pandemic = case_when(year %in% 2020:2021 ~ "pandemic",
                              TRUE ~ "the before times")) %>%
  ggplot(aes(month, avg, group = year, color = pandemic)) +
  geom_line(size = 1.4, alpha = 0.7) +
  scale_color_brewer(name = NULL,
                    direction = -1,
                    palette = "BrBG") +
  labs(
    x = NULL,
    color = NULL,
    y = "Median concurrent players for all games",
    title = "Video games on Steam and the pandemic",
    subtitle = "The overall median number of concurrent players is higher during the pandemic",
    caption = "Source: Kaggle"
  ) +
  theme(legend.position = "top")
```




