---
title: "Animating a geographic introduction"
date: "March 7, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup}
suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)
library(gganimate)
library(ggtext)
library(glue)
})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))

```

I was inspired by Sharon Pileggi's Rladies introdution, where she prepared a gif to introduce herself on Twitter for her week as host of the handle, as described at this [post](https://www.pipinghotdata.com/posts/2021-02-15-gganimating-a-geographic-introduction/).  I will follow a similar workflow, but for my own history and with my own choices of font and theme.

```{r}
us_states <- ggplot2::map_data("state") %>% 
  filter(region %in% c("iowa",  "wisconsin", "illinois",
                       "indiana","michigan"))
  
residences <- tribble(
  ~city, ~state, ~long, ~lat, ~years, ~description,
  "Lamont", "IA", -91.6476, 42.59999, 18, "Childhood",
  "Ames", "IA", -93.62018, 42.026802, 4, "Engineering Undergrad",
  "Muscatine", "IA", -91.04485, 41.42045, 2, "HON Chairs",
  "Racine", "WI", -87.791916, 42.72062, 5, "Case Corporation<br>Magnum Engineering",
  "Kenosha", "WI", -87.821185, 42.5847425, 2, "UW Parkside MBA",
  "Racine", "WI", -87.791916, 42.72062, 6, "CNH Industrial<br>Strategic Sourcing",
  "Burr Ridge", "IL", -87.9155, 41.74993, 8, "CNH Industrial<br>4WD Tractor",
  "Chicago", "IL", -87.623177, 41.881832, 1, "Notre Dame Mendoza",
  "Burr Ridge", "IL", -87.9155, 41.74993, 3, "CNH Industrial<br>Digital Operations",
)
```

Let's create a transition state for gganimate (city_order) as well as indicate connections between residences for the arrows.

```{r}
residence_connections_prelim <- residences %>% 
  mutate(
    # need this to create transition state ----
    city_order = row_number() + 1,
    # where I moved to next, for curved arrows ----
    lat_next = lead(lat),
    long_next = lead(long),
    # label to show in plot, styled using ggtext ---
    label = glue::glue("**{city}, {state}** ({years} yrs)<br>*{description}*"),
    # label of next location ----
    label_next = lead(label)
  ) 
```

Let's modify this data a bit so that the first residence shows the label at the residence with no arrow and all remaining residences show an arrow with the label positioned at the next residence.

```{r}
residence_connections <- residence_connections_prelim %>%
  # get first row of residence ----
  slice(1) %>% 
  # manually modify for plotting ----
  mutate(
    city_order = 1,
    label_next = label,
    lat_next = lat,
    long_next = long,
    ) %>% 
  # combine with all other residences ----
  bind_rows(residence_connections_prelim) %>% 
  # last (9th) row irrelevant ----
  slice(1:9) %>% 
  # keep what we neeed ----
  dplyr::select(city_order, lat, long, lat_next, long_next, label_next)

residence_connections
```

## Base map

The static base map is drawn in light gray with white state borders. Each marker is a line segment plus a point in a sort of lollipop.

```{r}
base_map <- ggplot() +
  geom_polygon(
    data = us_states,
    aes(
      x     = long, 
      y     = lat, 
      group = group
      ),
    fill  = "gray80",
    color = "white"
  ) +
  geom_segment(
    data = residences,
    aes(
      x    = long,
      xend = long,
      y    = lat,
      yend = lat + 0.2
      ),
    color = "midnightblue",
    size = 0.3
    ) +
  geom_point(
    data = residences,
    aes(
      x = long, 
      y = lat + 0.2
      ),
    size = 0.5,
    color = "red"
  ) +
#  coord_map() +
  coord_fixed() +
#  coord_sf(label_axes = "----") +
  theme(panel.grid.major = element_blank()) +
  labs(x = NULL, y = NULL) +
  theme_void()

base_map
```

## The animation

The two layers of plot elements in the animation draw richtext labels and curved arrows in sequence.

```{r}
anim <- base_map +
  # add in labels for pins, with inward positioning ----
  # show labels either top left or top right of pin ----
  geom_richtext(
    data = residence_connections,
    aes(
      x     = ifelse(long_next < -92, long_next + 1, 
                     long_next - 1),
      y     = ifelse(lat_next < 42, lat_next - 1.5,
                     lat_next + 3),
      label = label_next,
      vjust = "top",
      hjust = ifelse(long_next < -92, 0, 1),
      # group is used to create the transition ----
      group = seq_along(city_order)
    ),
    size = 2,
    label.colour = "white",
    color = "midnightblue",
  ) +
  # show arrows connecting residences ----
  geom_curve(
    # do not include 1st residence in arrows as no arrow is intended ----
    # and inclusion messes up transition ---
    data = residence_connections %>% slice(-1),
    # add slight adjustment to arrow positioning ----
    aes(
      y     = lat - 0.1,
      x     = long,
      yend  = lat_next - 0.1,
      xend  = long_next,
      # group is used to create the transition ----
      group = seq_along(city_order)
    ),
    color = "midnightblue",
    curvature = -0.5,
    arrow = arrow(length = unit(0.01, "npc")),
    size  = 0.2
  ) +
  # create animation ----
  transition_states(
    city_order,
    transition_length = 2,
    state_length = 10
    ) 
```

The options for gganimate that work for me include the use of the new `ragg_png` device and the `magick_renderer`.

```{r animation}
animation <- animate(anim, 
                     nframes = 30, 
                     height = 2, 
                     width = 3, 
                     duration = 60,
                     units = "in",
                     device = "ragg_png",
                     res = 150,
                     renderer = magick_renderer())

```


