---
title: "torch, tidymodels, and high-energy physics"
date: "February 11, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

I am going to make a departure from the TidyTuesday script this week to explore a Sigrid Keydana article posted at the R-Studio AI Blog as "[torch, tidymodels, and high-energy physics](https://blogs.rstudio.com/ai/posts/2021-02-11-tabnet/)," where they introduce [`tabnet`](https://mlverse.github.io/tabnet/), a torch implementation of the paper “[TabNet: Attentive Interpretable Tabular Learning](http://arxiv.org/abs/1908.07442)” that is fully integrated with the `tidymodels` machine learning framework. 

To showcase `tabNet`, Sigrid uses the [Higgs dataset (Baldi, Sadowski, and Whiteson (2014))](https://archive.ics.uci.edu/ml/datasets/HIGGS), available at the UCI Machine Learning Repository. The file is 2.6 Gb, zipped.

What caught my eye in her post was the claim that TabNet includes interpretability features by design. Let's give it a try here. First, let's load package libraries to make the deep learning functions available.

```{r setup}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	cache.lazy = FALSE,
	df_print = "paged",
	dpi = 300,
	tidy = "styler",
	dev = "ragg_png"
)

suppressPackageStartupMessages({
library(tidyverse)
library(torch)
library(tabnet)
library(tidymodels)
library(vip)
library(finetune)

library(rlang)
library(vctrs)

library(hrbrthemes)
library(systemfonts)
 })

ggplot2::theme_set(hrbrthemes::theme_ipsum_pub() +
            ggplot2::theme(plot.title.position = "plot",
                           plot.caption.position = "plot"))

```

I've invested some time and effort into making GPU compute available on my Windows laptop.

```{r}
cuda_device_count()
torch_get_num_interop_threads()
torch_get_num_threads()
```

Next, let's load the dataset into memory:

```{r load}
higgs <- read_csv(here::here("data", "HIGGS.csv"),
  col_names = c("class", 
                "lepton_pT", 
                "lepton_eta", 
                "lepton_phi", 
                "missing_energy_magnitude",
                "missing_energy_phi", 
                "jet_1_pt", 
                "jet_1_eta", 
                "jet_1_phi", 
                "jet_1_b_tag",
                "jet_2_pt", 
                "jet_2_eta", 
                "jet_2_phi", 
                "jet_2_b_tag", 
                "jet_3_pt", 
                "jet_3_eta",
                "jet_3_phi", 
                "jet_3_b_tag", 
                "jet_4_pt", 
                "jet_4_eta", 
                "jet_4_phi", 
                "jet_4_b_tag",
                "m_jj", 
                "m_jjj", 
                "m_lv", 
                "m_jlv", 
                "m_bb", 
                "m_wbb", 
                "m_wwbb"),
  col_types = "fdddddddddddddddddddddddddddd"
  )

levels(higgs$class) <- c(TRUE, FALSE)  # 1 is TRUE, 0 is FALSE

```

In high-energy physics, the search for new particles takes place at particle accelerators like Fermilab and CERN’s Large Hadron Collider. In addition to the actual experiments, simulation plays a role, synthesized for different underlying hypotheses, resulting in distributions that can be compared.

The `higgs` dataset (Baldi, Sadowski, and Whiteson (2014)) results from this sort of synthesized simulation. It explores features assuming two different processes. In the first, two gluons collide, and a heavy Higgs boson is produced; this is the signal process, the one we’re interested in. In the second, the collision of the gluons results in a pair of top quarks – this is the background process.

Through different intermediaries, both processes result in the same end products. What the paper authors did was simulate kinematic features of decay products, such as leptons (electrons and protons) and particle jets. In addition they constructed a number of high-level features that presuppose domain knowledge. In the article the show that, in contrast to other machine learning methods, deep neural networks did nearly as well when presented with the low-level features (the momenta) only as with just the high-level features alone.

Certainly, it would be interesting to double-check these results on `tabnet`, and then, look at the respective feature importance. Given the size of the dataset, some significant computing resources could be required.

Let’s take a glimpse:

```{r}
higgs %>% 
  glimpse()
```

```{r}
prop.table(table(higgs$class))

```

The first variable, class, is either TRUE or FALSE, depending on whether a Higgs boson was present or not. While in experiments only a tiny fraction of collisions produce one, both classes are about equally frequent in this dataset.

Within the predictors, the last seven are high-level (derived). All others are “measured”.

Let's split the data:

```{r}
n_all <- nrow(higgs)
n_test <- 500000
test_frac <- n_test / n_all # cleaned up compared to vignette

split <- initial_split(higgs, prop = 1 - test_frac, strata = class) # vignette used initial_time_split
train <- training(split)
test  <- testing(split)
```

And then create a `recipe`:

```{r}
rec <- recipe(class ~ ., train) 
```

Next, we will create a `parsnip` model specification of class `tabnet`. The parameters passed are those reported by the TabNet paper, for the S-sized model variant used on this dataset (except for the number of epochs).

```{r}

mod <- tabnet(epochs = 3,
              batch_size = 16384,
              decision_width = 24,
              attention_width = 26,
              num_steps = 5,
              penalty = 0.000001,
              virtual_batch_size = 512,
              momentum = 0.6,
              feature_reusage = 1.5,
              learn_rate = 0.02) %>% 
  set_engine("torch", verbose = TRUE) %>%
  set_mode("classification")
```

Next, lets bundle the recipe and model into a workflow for training

```{r}
wf <- workflow() %>%
  add_model(mod) %>%
  add_recipe(rec)
```

Next, train the model. This takes some time, even with the GPU. Training finished, let's save the trained parsnip model so we can reuse it without waiting every time we knit.

```{r initial_fit, eval = FALSE}
fitted_model <- wf %>% fit(data = train)

fitted_model$fit$fit$fit %>% saveRDS(here::here("data","saved_tabnet_model1.rds"))
```

```{r initial_fit_run, include = FALSE}
if (!file.exists(here::here("data","saved_tabnet_model1.rds"))) {
fitted_model <- wf %>% fit(train)
fitted_model$fit$fit$fit %>% saveRDS(here::here("data","saved_tabnet_model1.rds"))
} else {
fitted_model <- read_rds(here::here("data","saved_tabnet_model1.rds"))
}
```

`[Epoch 001] Loss: 0.851965`                               
`[Epoch 002] Loss: 0.637281`                               
`[Epoch 003] Loss: 0.605421  `

After just three epochs, my loss was 0.605. My model run time was `r fitted_model$fit$fit$elapsed[[3]]/60/60` hours, and the saved model file is `r file.info(here::here("data","saved_tabnet_model1.rds"))$size/1e6` Mb.

Finally, we ask the model for test-set predictions and compute performance metrics.

```{r}
metrics <- metric_set(accuracy, precision, recall)

bind_cols(test, predict(fitted_model, test)) %>% 
  metrics(class, estimate = .pred_class)
```

We didn’t quite arrive at the accuracy reported in the TabNet paper (0.783), but then, we only trained for a tiny fraction of the time.

Let's see how much we can improve with some hyperparameter tuning.

## Tabnet tuning

For hyperparameter tuning, the `tidymodels` framework makes use of cross-validation. With a dataset of considerable size, some time and patience is needed; for the purpose of this post, we will use 1/1,000 of the observations.

Changes to the above workflow start at model specification. Let’s say we’ll leave most settings fixed, but vary the TabNet-specific hyperparameters decision_width, attention_width, and num_steps, as well as the learning rate:

```{r}
mod <- tabnet(epochs = 1, 
              batch_size = 16384, 
              decision_width = tune(), 
              attention_width = tune(),
              num_steps = tune(), 
              penalty = 0.000001, 
              virtual_batch_size = 512, 
              momentum = 0.6,
              feature_reusage = 1.5, 
              learn_rate = tune()) %>%
  set_engine("torch", verbose = TRUE) %>%
  set_mode("classification")
```

The workflow is the same as before:

```{r}
wf <- workflow() %>%
  add_model(mod) %>%
  add_recipe(rec)
```

Next, we specify the hyperparameter ranges we’re interested in, and call one of the grid construction functions from the `dials` package to build one for us. 

```{r}
grid <- wf %>%
  parameters() %>%
  update(
    decision_width = decision_width(range = c(20, 30)),
    attention_width = attention_width(range = c(30, 40)),
    num_steps = num_steps(range = c(4, 6)),
    learn_rate = learn_rate(range = c(-2.5, -0.8),
                            trans = log10_trans())
  ) %>%
  grid_max_entropy(size = 8)

grid
```

To search the space, we use `tune_race_anova()` from the new `finetune` package, making use of five-fold cross-validation:

```{r}
ctrl <- control_race(verbose_elim = TRUE)
folds <- vfold_cv(train, v = 5)

res <- wf %>% 
    tune_race_anova(
    resamples = folds, 
    grid = grid,
    control = ctrl
  )
```

