---
title: "Sliced Bank Account Churn"
author: "Jim Gruman"
date: "July 13, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
    code_download: true
    df_print: paged
editor_options:
  chunk_output_type: console
---

[SLICED](https://www.notion.so/SLICED-Show-c7bd26356e3a42279e2dfbafb0480073) is like the TV Show Chopped but for data science. Competitors get a never-before-seen dataset and two-hours to code a solution to a prediction challenge. Contestants get points for the best model plus bonus points for data visualization, votes from the audience, and more.
 
[Season 1 Episode 7](https://www.kaggle.com/c/sliced-s01e07-HmPsw2/data) features a challenge to predict whether a bank customer is churned (lost to another bank). The evaluation metric for submissions in this competition is residual mean log loss.

To make the best use of the resources that we have, we will explore the data set for features to select those with the most predictive power, build a random forest to confirm the recipe, and then use the rest of the time to build one or more ensemble models. 

Let's load up some packages:

```{r setup}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	cache.lazy = FALSE,
	df_print = "paged",
	dpi = 300,
	tidy = "styler",
	dev = "ragg_png",
	autodep = TRUE
)
suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)
library(lubridate)

library(tidymodels)
library(treesnip)
library(finetune)

library(themis)
library(baguette)
  
library(catboost)

})

update_geom_font_defaults(font_rc_light)

ggplot2::theme_set(hrbrthemes::theme_ipsum_tw() +
            ggplot2::theme(plot.title.position = "plot",
                           plot.caption.position = "plot"))

#create a data directory
data_dir <- here::here("data",Sys.Date())
if (!file.exists(data_dir)) dir.create(data_dir)

# set a competition metric
mset <- metric_set(mn_log_loss)

# set the competition name from the web address
competition_name <- "sliced-s01e07-HmPsw2"

zipfile <- paste0(data_dir,"/", competition_name, ".zip")

path_export <- here::here("data",Sys.Date(),paste0(competition_name,".csv"))
```

# Import and Exploratory Data Analysis {.tabset}

A quick reminder before downloading the dataset:  Go to the web site and accept the competition terms!!!

## Direct Import and Shell Commands

We have basic shell commands available to interact with Kaggle here:

```{r kaggle competitions terminal commands, eval=FALSE}
# from the Kaggle api https://github.com/Kaggle/kaggle-api

# the leaderboard
shell(glue::glue('kaggle competitions leaderboard { competition_name } -s'))

# the files to download
shell(glue::glue('kaggle competitions files -c { competition_name }'))

# the command to download files
shell(glue::glue('kaggle competitions download -c { competition_name } -p { data_dir }'))

# unzip the files received
shell(glue::glue('unzip { zipfile } -d { data_dir }'))

```

Reading in the contents of the datafiles here:

```{r}
train_df <- read_csv(file = glue::glue({ data_dir }, "/train.csv")) %>% 
  mutate(across(c(attrition_flag, gender, education_level, income_category), as.factor)) 

test_df <- read_csv(file = glue::glue({ data_dir }, "/test.csv")) %>% 
  mutate(across(c(gender, education_level, income_category), as.factor)) 

```

## Skim

Some questions to answer here:
What features have missing data, and imputations may be required?
What does the outcome variable look like, in terms of imbalance?

```{r}
skimr::skim(train_df)
```

Outcome variable `attrition_flag` is set to 1 (churned) for 1132 rows, and 0 (not churned) for 5956, so is imbalanced. None of the features is missing obvious data. We will take a closer look at the categorical variable levels in a moment.

## Categorical Feature Plots

```{r}
summarize_attrition <- function(tbl) {
  ret <- tbl %>%
    summarize(
      n_attrition = sum(attrition_flag == 1),
      n = n(),
      .groups = "drop"
    ) %>%
    arrange(desc(n)) %>%
    mutate(
      pct_attrition = n_attrition / n,
      low = qbeta(.025, n_attrition + 5, n - n_attrition + .5),
      high = qbeta(.975, n_attrition + 5, n - n_attrition + .5)
    ) %>%
    mutate(pct = n / sum(n))
  ret
}

train_df %>%
  summarize_attrition()

train_df %>%
  group_by(education_level) %>%
  summarize_attrition() %>%
  mutate(education_level = fct_reorder(education_level, pct_attrition)) %>%
  ggplot(aes(pct_attrition, education_level)) +
  geom_point(aes(size = pct)) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = .3) +
  scale_size_continuous(labels = percent,
                        guide = "none",
                        range = c(.5, 4)) +
  scale_x_continuous(labels = percent) +
  labs(
    x = "Proportion of attrition",
    y = "",
    title = "What education levels get the most attrition?",
    subtitle = "Including 95% intervals. Size of points is proportional to frequency in the dataset"
  )

train_df %>%
  group_by(income_category = fct_lump(income_category , 30)) %>%
  summarize_attrition() %>%
  mutate(income_category = fct_reorder(income_category, pct_attrition)) %>%
  ggplot(aes(pct_attrition, income_category)) +
  geom_point(aes(size = pct)) +
  geom_errorbarh(aes(xmin = low, xmax = high), height = .3) +
  scale_size_continuous(labels = percent,
                        guide = "none",
                        range = c(.5, 4)) +
  scale_x_continuous(labels = percent) +
  labs(
    x = "Proportion of attrition",
    y = "",
    title = "What income category levels get the most attrition?",
    subtitle = "Including 95% confidence intervals. Size of points is proportional to frequency in the dataset"
  )

train_df %>%
  ggplot(aes(gender, credit_limit, color = attrition_flag)) +
  geom_boxplot(show.legend = FALSE) +
  theme(
    plot.subtitle = ggtext::element_textbox_simple(),
    plot.background = element_rect(color = "white")
  ) +
  labs(
    title = "Differences in Credit Limits, by Gender",
    subtitle = "Customers <span style = 'color:#00BFC4'>churning</span> and <span style = 'color:#F8766D'> not churning</span>. Note that the vast majority of Men have higher limits.",
    x = "Gender",
    y = "Credit Limit"
  )

```

## Numeric Feature Plots

The outcome variable itself is skewed across all observations in the training data, as prices often are.

```{r}
train_numeric <- train_df %>% keep(is.numeric) %>% colnames()

chart <- c(train_numeric, "attrition_flag")

p1 <- train_df %>%
  select_at(vars(chart)) %>%
  select(-id) %>%
  pivot_longer(
    cols = -attrition_flag,
    names_to = "key",
    values_to = "value"
  ) %>%
  filter(!is.na(value)) %>%
  ggplot(mapping = aes(value, 
  #                     after_stat(density), 
                       fill = attrition_flag)) +
  geom_histogram(
    position = "identity",
    alpha = 0.5,
    bins = 30,
    show.legend = FALSE
  ) +
  facet_wrap( ~ key, scales = "free", ncol = 3) +
  scale_x_continuous(n.breaks = 3) +
  theme(
    plot.subtitle = ggtext::element_textbox_simple(),
    plot.background = element_rect(color = "white")
  ) +
  labs(
    title = "Numeric Feature Histogram Distributions",
    subtitle = "Customers <span style = 'color:#00BFC4'>churning</span> and <span style = 'color:#F8766D'> not churning</span>",
    x = "Numeric Feature",
    y = NULL
  )

ggsave(p1, height = 7, width = 7,
       file = here::here("docs", "assets", "attrition.png"),
       device = "png")
```

```{r}
knitr::include_graphics(here::here("docs", 
                                   "assets",
                                   "attrition.png"),
                         error = FALSE)
```

## Numeric Feature Correlations

```{r}
cr1 <- train_df %>%
  keep(is.numeric) %>%
  cor(use = "pair")

corrplot::corrplot(cr1, type = "upper")

```

The transaction amounts and transaction counts tend to be correlated. Also, the revolving balance and average utilization ratio.

# {-}

----

Conclusions:

* There aren't any obvious features to engineer, so as a first pass we will go straight to training hyperparameters with the features we have.
* As a second pass, we can add interactions to try to improve the model.

# Preprocessiong {.tabset}

## The recipe framework

I am going to train on almost all of the train data. The 1% left is a quick confirmation of validity.  The provided file labeled `test` has no labels.

```{r}
set.seed(2021)
split <- train_df %>% 
  initial_split(strata = attrition_flag,
                       prop = .99)
train <- training(split)
valid <- testing(split)
```

There are only `r nrow(valid)` held out from training as a last check before submission.

```{r}
basic_rec <-
  recipe(attrition_flag ~ ., train) %>%
  # ---- set aside the row id's
  update_role(id, new_role = "id") %>% 
  step_novel(gender, education_level, income_category) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_nzv(all_predictors())
```

## the pre-processed data

```{r}
basic_rec %>% 
#  finalize_recipe(list(num_comp = 2)) %>% 
  prep() %>% 
  juice() 
```

## Cross Validation

We will use 5-fold cross validation and stratify between the rain and no-rain classes.

```{r}
train_folds <- vfold_cv(data = train,
                        strata = attrition_flag,
                        v = 5)

```

# {-}

# Machine Learning {.tabset}

Let's run models in two steps. The first is a simple, fast shallow random forest, to confirm that the model will run and observe feature importance scores. The second will use `catboost`. Both use the basic recipe preprocessor.

## Model Specifications

We will build a specification for simple shallow random forest and a specification for catboost. 

```{r}
catboost_spec <- boost_tree(trees = 1000,
                            min_n = tune(),
                            learn_rate = tune(),
                            tree_depth = tune()) %>% 
  set_engine("catboost") %>%
  set_mode("classification")

bag_spec <-
  bag_tree(min_n = 10) %>%
  set_engine("rpart", times = 50) %>%
  set_mode("classification")

```

## Parallel backend

To speed up computation we will use a parallel backend.

```{r}
all_cores <- parallelly::availableCores(omit = 1)
all_cores

future::plan("multisession", workers = all_cores) # on Windows

```

## A quick Random Forest

Lets make a cursory check of the recipe and variable importance, which comes out of `rpart` for free. This workflow also handles factors without dummies.

```{r}
bag_wf <-
  workflow() %>%
  add_recipe(basic_rec) %>%
  add_model(bag_spec)

set.seed(123)
bag_fit <- parsnip::fit(bag_wf, data = train)

pull_workflow_fit(bag_fit)$fit$imp %>%
  mutate(term = fct_reorder(term, value)) %>%
  ggplot(aes(value, term)) +
  geom_point() +
  geom_errorbarh(aes(
    xmin = value - `std.error` / 2,
    xmax = value + `std.error` / 2
  ),
  height = .3) +
  labs(title = "Feature Importance",
       x = NULL, y = NULL)

```

We see that `total_trans_ct` and `total_trans_amt` will be very important for this model.

## Tuning Catboost

Now that we have some confidence that the features have predictive power, lets tune up a set of `catboost` models. 

```{r catboost_tuning}

catboost_params <-
  dials::parameters(min_n(), # min data in leaf
                    tree_depth(range = c(4, 15)),
                    learn_rate(range = c(-3, -0.7), 
                               trans = log10_trans())
                    )
                    
cbst_grid <- dials::grid_max_entropy(catboost_params,
                                     size = 40 
                                     )
cbst_grid

cv_res_catboost <-
  workflow() %>% 
  add_recipe(basic_rec) %>% 
  add_model(catboost_spec) %>% 
  tune_grid(    
    resamples = train_folds,
    grid = cbst_grid,
    control = control_race(verbose = FALSE,
                           save_pred = TRUE, 
                           save_workflow = TRUE,
                           extract = extract_model,
                           parallel_over = "resamples"),
    metrics = mset
)

autoplot(cv_res_catboost)

show_best(cv_res_catboost) %>% 
  select(-.estimator)

cat_wf_best <-   
  workflow() %>% 
  add_recipe(basic_rec) %>% 
  add_model(catboost_spec) %>% 
  finalize_workflow(select_best(cv_res_catboost))

cat_fit_best <- cat_wf_best %>%
  parsnip::fit(data = train)

```

# {-}

## Catboost model performance

```{r}
predict(cat_fit_best, new_data = valid, type = "prob") %>% 
  cbind(valid) %>%
  mn_log_loss(attrition_flag , `.pred_0`)

```

This catboost mean logloss figure is not bad, so I'm going to write it out in a suitable csv...

```{r}

bind_cols(predict(cat_fit_best, test_df), test_df) %>% 
  select(id, attrition_flag = `.pred_class`) %>%
  write_csv(file = path_export)

```

and make the submission to the Kaggle board

```{r, eval = FALSE}
shell(glue::glue('kaggle competitions submit -c { competition_name } -f { path_export } -m "Catboosted with numeric interactions"'))
```

My interim standings on the competition leaderboard:

```{r}
knitr::include_graphics(here::here("docs", 
                                   "assets",
                                   "2021-07-13.png"),
                         error = FALSE)
```

The standings will likely change after Kaggle rolls the scores for the full test dataset.

## Machine Learning Round 2 {.tabset}

## Recipe with Interactions

Now modeling with the numeric interaction features, tightening up the learn rate, and reducing the size of the grid to speed up results. We are nearly out of time.

```{r advanced_recipe}

advanced_rec <-
  recipe(attrition_flag ~ ., train) %>%
  # ---- set aside the row id's
  update_role(id, new_role = "id") %>% 
  step_novel(gender, education_level, income_category) %>%
  step_interact(terms = ~ total_relationship_count:avg_utilization_ratio) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_nzv(all_predictors())
```

## Advanced Catboost Tuning

```{r advanced_catboost_tuning}

catboost_params <-
  dials::parameters(min_n(), # min data in leaf
                    tree_depth(range = c(4, 15)),
                    learn_rate(range = c(-1, -0.7), # updated
                               trans = log10_trans())
                    )
                    
cbst_grid <- dials::grid_max_entropy(catboost_params,
                                     size = 20
                                     )
cbst_grid

cv_res_catboost <-
  workflow() %>% 
  add_recipe(advanced_rec) %>% 
  add_model(catboost_spec) %>% 
  tune_grid(    
    resamples = train_folds,
    grid = cbst_grid,
    control = control_race(verbose = FALSE,
                           save_pred = TRUE, 
                           save_workflow = TRUE,
                           extract = extract_model,
                           parallel_over = "resamples"),
    metrics = mset
)

autoplot(cv_res_catboost)

show_best(cv_res_catboost) %>% 
  select(-.estimator)

cat_wf_best <-   
  workflow() %>% 
  add_recipe(basic_rec) %>% 
  add_model(catboost_spec) %>% 
  finalize_workflow(select_best(cv_res_catboost))

cat_fit_best <- cat_wf_best %>%
  parsnip::fit(data = train)

```

## Advanced catboost model performance

```{r}
predict(cat_fit_best, new_data = valid, type = "prob") %>% 
  cbind(valid) %>%
  mn_log_loss(attrition_flag , `.pred_0`)

```

This catboost figure is not an improvement, so I'm not going to submit.

# {-}

----

My final standings on the Kaggle competition leaderboard:

```{r}
knitr::include_graphics(here::here("docs", 
                                   "assets",
                                   "2021-07-13-2.png"),
                         error = FALSE)
```


