---
title: "Eurovision Grand Finale Rankings"
author: "Jim Gruman"
date: "May 17, 2022"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
    code_download: true
    df_print: paged
editor_options:
  chunk_output_type: console
---

For #TidyTuesday 2022 week 20, Bob Rudis and Tanya Shapiro built a nice sample dataset of Eurovision grand finale rankings from 2004 to 2022.

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  fig.width = 12, cache = FALSE, warning = FALSE,
  fig.height = 9,
  echo = TRUE, cache.lazy = FALSE, tidy = "styler"
)

```

I am going to attempt to build on Lee Olney's submission to twitter. Reference [https://github.com/leeolney3/TidyTuesday/blob/main/2022/week_20/](https://github.com/leeolney3/TidyTuesday/blob/main/2022/week_20/)

```{r}
#| label: tweet-inspiration
#| echo: false

tweetrmd::include_tweet("https://twitter.com/leeolney3/status/1526448720392466437")
```

First, let's load the packages:

```{r}
#| label: load-packages

suppressPackageStartupMessages({
library(tidyverse)
library(cowplot)
})

```

Second, let's load the data and a helper function to glue the counts of countries onto the country labels:

```{r}
#| label: load_tidytuesdayR_data
withfreq <- function(string, width = 20) {
  dplyr::pull(dplyr::mutate(
    dplyr::add_count(tibble::tibble(string), string),
    string = glue::glue("{ stringr::str_wrap(string, width = width) } ({ n })")
  ),
  string)
}

eurovision <- tidytuesdayR::tt_load("2022-05-17")$eurovision |> mutate(artist_country = withfreq(artist_country))

# h/t to Priyanka Mehta @Priyank79286307 for the suggestion to order country by which countries participated most and which won the most

eurovision_by_country <- eurovision |>
  filter(section == "grand-final", year != 2020) |>
  group_by(year) |>
  mutate(rank_label = case_when(
    rank %in% c(1, 2, 3) ~ as.character(rank),
    rank == max(rank) ~ "last",
    TRUE ~ NA_character_
  )) |>
  ungroup(year) |> 
  mutate(
    artist_country = factor(artist_country, 
          levels = 
            eurovision |>
              group_by(artist_country)  |>
              summarise(highest_rank = min(rank),
                      n = n_distinct(year),
                      .groups = "drop")  |> 
              arrange(n, desc(highest_rank)) |> 
              pull(artist_country))
    )


```

```{r}
#| plot

eurovision_by_country |> 
  ggplot(aes(x = year, y =	artist_country)) +
  geom_line( size = .3, color = "grey70") +
  geom_point(shape = 21, size = 2.5, fill = "white") +
  geom_point(
    data = eurovision_by_country |> filter(!is.na(rank_label)),
    aes(fill = rank_label),
    size = 2.5,
    shape = 21
  ) +
  scale_fill_manual(values = c("#F50405", "#F7C83A", "#1CB4EB","#4F5251"),
                    guide = guide_legend(order = 1)) +
  scale_x_continuous(
    position = "top",
    breaks = seq(2005, 2020, 5),
    expand = expansion(mult = c(.02, NA))
  ) +
  coord_cartesian(clip = "off") +
  cowplot::theme_minimal_vgrid(12) +
  theme(
    legend.title = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "top",
    plot.margin = margin(.4, 1.4, .3, .4, unit = "cm"),
    plot.title.position = "plot",
    plot.title = element_text(size = 13),
    plot.caption.position = "plot",
    plot.caption = element_text(
      hjust = 0,
      color = "grey20",
      margin = margin(t = 13),
      size = 9
    )
  ) +
  labs(title = "Eurovision Grand Final Rankings",
       subtitle = "From 2004 to 2022, arranged in order of artist country number of appearances (in parens)",
       caption = "#TidyTuesday week 20 | Data from Eurovision, credits to Tanya Shapiro and Bob Rudis and Lee Olney")
 
```

Thank you to 

- Lee Olney for the head start
- Bob Rudis and Tanya Shapiro


