---
title: "Animal Crossing Sentiment Analysis"
author: "Jim Gruman"
date: "May 5, 2020"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
  html_document:
    toc: no
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup}
suppressPackageStartupMessages({
library(tidyverse)
library(tidytext)
library(lubridate)
library(tidyr)
library(stm)
library(tidymodels)
library(textrecipes)
library(vip)
})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())
ggplot2::theme_set(theme_jim(base_size = 12))
```

Our goal this week is to explore and predict ratings for [Animal Crossing user reviews in this week’s #TidyTuesday dataset](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-05/readme.md) from the text in the review. This is what is typically called a sentiment analysis modeling, and it’s a common real-world problem. 

## Explore the data

```{r}

user_reviews <- readr::read_tsv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/user_reviews.tsv")

user_reviews %>%
  count(grade) %>%
  ggplot(aes(factor(grade), n)) +
  geom_col(aes(fill = factor(grade)),
           show.legend = FALSE) +
  labs(
    title = "Most Animal Crossing Reviews are at the Extremes",
    subtitle = paste0(
      "User Review Scores from ",
      format(min(user_reviews$date), "%B %d, %Y")
      ,
      " to ",
      format(max(user_reviews$date), "%B %d, %Y")
    ),
    caption = "Jim Gruman | #TidyTuesday" ,
    x = "Grade",
    y = "Number of Reviews"
  )
```

This phenomenon is common in surveys, where many people give the extreme scores of zero and 10. This does not look like a nice distribution for predicting a not-really-continuous quantity like grade, so we’ll convert these user scores to a label, good vs. bad user reviews, and build a classification model. A common theme for the negative reviews is frustration with the one-island-per-console setup, and more specifically the relative roles of player 1 vs. others on the same console.

There is definitely evidence of scraping problems when looking at the review text. Let’s remove at least the final "Expand" from the reviews, and create a new categorical rating variable.

```{r}
reviews_parsed <- user_reviews %>%
  mutate(text = str_remove(text, "Expand$")) %>%
  mutate(rating = factor(case_when(
    grade > 7 ~ "good",
    TRUE ~ "bad"
  )))
```

What is the distribution of the new, dependent, category variable **rating**?

There is a slight skew but we will soldier on and ignore it. 

```{r}
reviews_parsed  %>%
  ggplot(aes(rating, fill = rating)) +
  geom_bar(show.legend = FALSE) +
  labs(title = "With a cutoff at 8-10 as Good, how balanced is the dataset?")
```

What is the distribution of words per review?

The graph suggests that something is driving a bimodal distribution, with a gap at about 100 words. It should be investigated further.

```{r}
reviews_parsed %>%
  unnest_tokens(word, text) %>%
  count(user_name, name = "total_words") %>%
  ggplot(aes(total_words)) +
  geom_histogram(fill = "#2AA198", bins = 30) +
  scale_x_log10() +
  labs(
    title = "Count of Words in Animal Crossing Reviews",
    subtitle = paste0(
      "Reviews from ",
      format(min(user_reviews$date), "%B %d, %Y")
      ,
      " to ",
      format(max(user_reviews$date), "%B %d, %Y")
    ),
    caption = "Jim Gruman | #TidyTuesday" ,
    x = "Number of Words in Review",
    y = "Number of Reviews"
  ) +
  geom_vline(xintercept = 100, color = "firebrick") +
  annotate(
    "text",
    x = 105,
    y = 325,
    hjust = 0,
    color = "firebrick",
    label = "Sharp cliff\nat 100 words"
  ) 
```

After a cursory inspection of the text of the reviews, there are some where the author pasted the same phrase into the field verbatim several times. We will filter out the repeats above 3x in the modeling dataset.

### Review Grades over time

Let's take a look at how the reviews evolve over time:

```{r}
by_week <- reviews_parsed %>%
  group_by(week = floor_date(date, "week", week_start = 1)) %>%
  summarize(nb_reviews = n(),
            avg_grade = mean(grade),
            pct_zero = mean(grade == 0),
            pct_ten = mean(grade == 10)) 

by_week %>% 
  ggplot(aes(week, avg_grade)) +
  geom_line(color = "#2AA198") +
  geom_point(aes(size = nb_reviews), color = "#2AA198") +
  expand_limits(y = 0) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  labs(x = "Week", y = "Average Grade",
       size = "# of Reviews",
       title = "Animal Crossing User Review Grades",
       subtitle = paste0("By Week"),
       caption = "Jim Gruman | #TidyTuesday" )

by_week %>% 
  gather(type, value, contains("pct")) %>%
  mutate(type = ifelse(type == "pct_zero", "Rated 0", "Rated 10")) %>%
  ggplot(aes(week, value, color = type)) +
  geom_line(show.legend = FALSE) +
  geom_point(aes(size = nb_reviews)) +
  expand_limits(y = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  labs(x = "Week", y = "Portion of Reviews",
       size = "# of Reviews",
       title = "Most Polarizing Animal Crossing Reviews",
       subtitle = paste0("By Week"),
       caption = "Jim Gruman | #TidyTuesday" )
```

Let's take a look at how the reviews evolve by day of the week:

```{r}
by_dow <- reviews_parsed %>%
  group_by(dow = wday(date, label = TRUE, week_start = 1)) %>%
  summarize(nb_reviews = n(),
            avg_grade = mean(grade),
            pct_zero = mean(grade == 0),
            pct_ten = mean(grade == 10)) 

by_dow %>% 
  ggplot(aes(dow, avg_grade)) +
  geom_point(aes(size = nb_reviews), color = "#2AA198") +
  expand_limits(y = 0) +
#  scale_x_date(date_labels = "%b %d", date_breaks = "1 week")+
  labs(x = "Day", y = "Average Grade",
       size = "# of Reviews",
       title = "Animal Crossing User Review Grades",
       subtitle = paste0("By Day of Week"),
       caption = "Jim Gruman | #TidyTuesday" )

by_dow %>% 
  gather(type, value, contains("pct")) %>%
  mutate(type = ifelse(type == "pct_zero", "Rated 0", "Rated 10")) %>%
  ggplot(aes(dow, value, color = type)) +
  geom_point(aes(size = nb_reviews), 
             shape = 21) +
  expand_limits(y = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Day", y = "Portion of Reviews",
       size = "# of Reviews",
       title = "Most Polarizing Animal Crossing Reviews",
       color = "Review Grade",
       subtitle = paste0("By The Day of the Week"),
       caption = "Jim Gruman | #TidyTuesday" )

```

For further topic modeling, we will remove the stop words and excessive copy-pasting within each review document.

```{r}
user_review_words <- reviews_parsed %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(user_name, word) %>%
  mutate(id = row_number()) %>%
  filter(id < 4) %>%
  ungroup() %>%
  count(user_name, date, grade, word)
```

What are the words that are positively or negatively associated with each user grade?

```{r}
by_word <- user_review_words %>%
  group_by(word) %>%
  summarize(avg_grade = mean(grade),
            nb_reviews = n()) %>%
  filter(nb_reviews >= 25) %>%
  arrange(desc(avg_grade)) 

by_word %>%
  filter(nb_reviews >= 75) %>%
  ggplot(aes(nb_reviews, avg_grade)) +
  geom_point() +
  geom_text(aes(label = word),
            vjust = 1, hjust = 1, 
            check_overlap = TRUE) +
  scale_x_log10()
```

To make a nice chart, let's zoom in on the 20 words associated with the lowest Review Grades:

```{r}
by_word %>%
  top_n(20, -avg_grade) %>%
  ggplot(aes(nb_reviews, avg_grade)) +
  geom_point() +
  geom_text(aes(label = word),
            vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_x_log10() +
  labs(x = "Number of Reviews", y = "Mean Review Grade",
       title = "Words Associated with Lowest Animal Crossing Reviews",
       subtitle = "20 most common words in unfavorable reviews; Appearing at least 25 times",
       caption = "Jim Gruman | #TidyTuesday" )
```

### Structural Topic Model 

The `stm` Structural Topic Model package offers unsupervised approaches to building clusters of topics, their relationships with documents, and their relationships with words.

We will arbitrarily choose to cluster here into 6 topics.

```{r}

review_matrix <- user_review_words %>%
  group_by(word) %>%
  filter(n() >= 25) %>%
  cast_sparse(user_name, word, n)

topic_model <- stm(review_matrix, 
                   K = 6,
                   verbose = FALSE, 
                   init.type = "Spectral"
                   )

```

We can plot the words most highly associated with each of 6 topics:

```{r} 
tidy(topic_model) %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ggplot(aes(beta, reorder_within(term, by = beta, within = topic),fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  tidytext::scale_y_reordered() +
  facet_wrap(~ topic, scales = "free_y") +
  labs(x = "beta",
       y = "Word Tokens",
       title = "Animal Crossing Review Topics",
       caption = "@jim_gruman | #TidyTuesday") +
  theme(axis.text.x.bottom = element_text(size = 8),
        plot.margin = unit(c(c(1, 1, 1, 0.5)), units = "cm"))
```

Is there a correlation between the document gamma and the grade that the reviewer gave? We will use the Spearman Correlation, which is sensitive to the ranks of the items.  And once again, we will examine how each topic is represented over time.

```{r}
topic_model_gamma <- tidy(topic_model, matrix = "gamma") %>%
  mutate(user_name = rownames(review_matrix)[document])  %>%
    inner_join(user_reviews, by = "user_name")

topic_model_gamma %>%
  group_by(topic) %>%
  summarize(spearman_correlation = cor(gamma, grade, method = "spearman"),
            .groups = "drop") %>%
  knitr::kable()
  
topic_model_gamma %>%
  group_by(week = floor_date(date, "week", week_start = 1),
           topic) %>%
  summarize(avg_gamma = mean(gamma),
            .groups = "drop") %>%
  ggplot(aes(week, avg_gamma, color = factor(topic))) +
  geom_line(size = 2, show.legend = FALSE) +
  geom_text(data = . %>% filter(week == as.Date("2020-04-27")),
            aes(label = paste0("Topic ", topic)),
            nudge_x = 2,
            check_overlap = TRUE,
            show.legend = FALSE) +
  expand_limits(y = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_date(date_labels = "%b %d", 
               date_breaks = "1 week",
               limits = c(NA_Date_, as.Date("2020-04-30"))) +
  labs(x = "Date", color = "Topic",
       y = "Average gamma (document-topic association)",
       title = "Animal Crossing Review Topic Evolution",
       caption = "@jim_gruman | #TidyTuesday")

```

## Build a Model

We start by loading the tidymodels metapackage, and splitting our data into training and testing sets.

```{r}

review_split <- initial_split(reviews_parsed, strata = rating)
review_train <- training(review_split)
review_test <- testing(review_split)
```

Next, let’s preprocess the data to get it ready for modeling. We can use specialized steps from textrecipes, along with the general recipe steps.

```{r}

review_rec <- recipe(rating ~ text + date, data = review_train) %>%
  step_mutate(date = as.numeric(difftime(date, min(date), units = "days"))) %>%
  step_mutate(dow = wday(date, label = TRUE, week_start = 1)) %>%
  step_dummy(dow) %>%
  step_tokenize(text) %>%
  step_stopwords(text) %>%
  step_tokenfilter(text, max_tokens = 500) %>%
  step_tfidf(text) %>%
  step_normalize(all_predictors())

review_prep <- prep(review_rec)

review_prep
```

Let’s walk through the steps in this recipe, which are sensible defaults for a first attempt at training a text classification model such as a sentiment analysis model.

- First, we must tell the `recipe()` what our model is going to be (using a formula here) and what data we are using.

- Next, we tokenize our text, with the default tokenization into single words.

- Next, we remove stop words (again, just the default set).

- It wouldn’t be practical to keep all the tokens from this whole dataset in our model, so we can filter down to only keep, in this case, the top 500 most-used tokens (after removing stop words). This is a pretty dramatic cut and keeping more tokens would be a good next step in improving this model.

- We need to decide on some kind of weighting for these tokens next, either something like term frequency or, what we used here, tf-idf.
`
- Finally, we center and scale (i.e. normalize) all the newly created tf-idf values because the model we are going to use is sensitive to this.

Before using `prep()` these steps have been defined but not actually run or implemented. The `prep()` function is where everything gets evaluated.

Now it’s time to specify our model. Here we can set up the model specification for lasso regression with `penalty = tune()` since we don’t yet know the best value for the regularization parameter and mixture = 1 for lasso. The lasso is often a good baseline for text modeling. 

I am using a `workflow()` in this example for convenience; these are objects that can help you manage modeling pipelines more easily, with pieces that fit together like Lego blocks. This `workflow()` contains both the recipe and the model.

```{r}
lasso_spec <- logistic_reg(penalty = tune(), 
                           mixture = 1) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(review_rec) %>%
  add_model(lasso_spec)

lasso_wf
```

### Tune model parameters

Let’s get ready to `tune` the lasso model! First, we need a set of possible regularization parameters to try.

```{r}
lambda_grid <- grid_regular(penalty(), levels = 40)
```

Next, we need a set of resampled data to fit and evaluate all these models.

```{r folds}
review_folds <- bootstraps(review_train, strata = rating)
```

Now we can put it all together and implement the tuning. We can set specific metrics to compute during tuning with `metric_set()`. Let’s look at AUC, positive predictive value, and negative predictive value so we can understand if one class is harder to predict than another.

```{r}
all_cores <- parallelly::availableCores(omit = 1)
all_cores

future::plan("multisession", workers = all_cores) # on Windows


lasso_grid <- tune::tune_grid(
  lasso_wf,
  resamples = review_folds,
  grid = lambda_grid,
  metrics = yardstick::metric_set(yardstick::roc_auc, yardstick::ppv, yardstick::npv)
)
```

Once we have our tuning results, we can examine them in detail. Visualization is often more helpful to understand model performance.

```{r lassoGrid}
lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  facet_wrap(~.metric) +
  scale_x_log10() +
  labs(title = "Model Performance at various Penalty Values")
```

This shows us a lot. We see clearly that AUC and PPV have benefited from the regularization and we could identify the best value of penalty for each of those metrics. The same is not true for NPV. One class (the happy comments) is harder to predict than the other. It might be worth including more tokens in our model, based on this plot.

### Choose the final model

Let’s keep our model as is for now, and choose a final model based on AUC. We can use `select_best()` to find the best AUC and then update our workflow lasso_wf with this value.

```{r}
best_auc <- lasso_grid %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  top_n(1, wt = mean) %>%
  select(penalty)

best_auc
```

```{r}
final_lasso <- finalize_workflow(lasso_wf, best_auc)

```

This is our tuned, finalized workflow (but it is not fit yet). One of the things we can do when we start to fit this finalized workflow on the whole training set is to see what the most important variables are using the `vip` package.

```{r finalLasso, fig.asp=1}
final_lasso %>%
  fit(review_train) %>%
  extract_fit_parsnip() %>% 
  vip::vi(lambda = best_auc$penalty) %>%
  group_by(Sign) %>%
  top_n(20, wt = abs(Importance)) %>%
  ungroup() %>%
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "tfidf_text_"),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~Sign, scales = "free_y") +
  labs(y = NULL) +
  labs(title = "Animal Crossing Review Word Importance",
       caption = "@jim_gruman | #TidyTuesday")
```

People who are happy with Animal Crossing like to talk about how relaxing, fantastic, enjoyable, and great it is, and also talk in their reviews about the “review bombing” of the negative reviews. Notice that many of the words from the negative reviews are specifically used to talk about the multiplayer experience. These users want a fix and they declare Nintendo greedy for the one-island-per-console play.

```{r variableImportance}
final_lasso %>%
  fit(review_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = best_auc$penalty) %>%
  group_by(Sign) %>%
  ungroup() %>%
  mutate(id = row_number()) %>%
  filter(str_detect(Variable, "dow")) %>%
  mutate(
    Importance = abs(Importance),
    Variable = str_remove(Variable, "dow_"),
    Variable = fct_reorder(Variable, Importance),
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col(show.legend = FALSE) +
  geom_label(aes(label = id), x = 0.1, show.legend = FALSE) +
  facet_wrap(~Sign, scales = "free_y") +
  labs(y = NULL) +
  labs(title = "Animal Crossing Review Day of Week Importance",
       caption = "@jim_gruman | #TidyTuesday")
```

The Day of Week feature does not appear in the Top 20. In this model, a Review written on a Wednesday (day 3) is 88th in importance, after 87 tokenized words. It has a positive influence. Thursday, day 4, is negative.

Finally, let’s return to our test data. The `tune` package has a function `last_fit()` which is nice for situations when you have tuned and finalized a model or workflow and want to fit it one last time on your training data and evaluate it on your testing data. You only have to pass this function your finalized model/workflow and your split.

```{r finalMetrics}
review_final <- last_fit(final_lasso, review_split)

review_final %>%
  collect_metrics() %>% 
  knitr::kable()
```

We did not overfit during our tuning process, and the overall accuracy is not bad. Let’s create a confusion matrix for the testing data.

```{r finalPredictions}
review_final %>%
  collect_predictions() %>%
  conf_mat(rating, .pred_class) 
```

Though our overall accuracy isn’t great, we discover here that it is easier to detect the negative reviews than the positive ones.

