---
title: "Beer Production"
author: "Jim Gruman"
date: 'March 31, 2020'
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
  html_document:
    toc: no
    df_print: paged
editor_options:
  chunk_output_type: console
name: beer_production
owner: jim_gruman
metrics:
  total_pounds:
    title: "# of pounds in production"
    description: "Number of pounds used in Beer Production in the US"
dimensions:
  material_type:
    title: Type
    description: Grain or nonGrain Products
  material:
    title: Material
    description: Wheat, hops, malt, barley, etc.
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(tidyverse)
library(lubridate)
library(tidymetrics)  #at github ramnathv/tidymetrics
library(GGally)
library(gghighlight)
library(tidymodels)
library(maps)
library(sf)
library(gganimate)
library(transformr)
})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))

```

## Import this week's data
```{r imports}
brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv') %>%
    mutate(date = ymd(paste0(year, "/", month, "/1"))) 
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
```

## Exploration

```{r, skims, eval = FALSE}
skimr::skim(beer_states)
# 19 year/states with missing barrels

skimr::skim(brewing_materials)
table(brewing_materials$type)
# opportunity to examine the proportional trends of Corn vresus Sugar/syrups use

skimr::skim(beer_taxed)

skimr::skim(brewer_size)
table(brewer_size$brewer_size)
```

Total all barrels produced by state per year (regardless of type)

```{r allPerStatePerYear}
all_barrels <-
  beer_states %>%
  na.omit() %>%
  filter(state != 'total') %>%
  group_by(state, year) %>%
  summarise(barrels = sum(barrels),
            .groups = "drop")
```

The top 5 states with largest increase in barrels from 2008 to 2019

```{r topFiveStates}
increase_states <-
  all_barrels %>%
  group_by(state) %>% 
  mutate(total_barrels = cumsum(barrels)) %>%
  filter(year %in% c(2008, 2019)) %>%
  mutate(pct_chg = (total_barrels - lag(total_barrels)) / lag(total_barrels)) %>%
  ungroup() %>%
  top_n(5, pct_chg) %>%
  pull(state)    

increase_states %>% 
  knitr::kable()
```

```{r makeBeerDf}
#Lets get the state names
states <- tibble(state = state.abb, state_name = state.name)

beer_df <- 
  beer_states %>%
  na.omit() %>%
  filter(state != 'total') %>%
  group_by(state, year) %>%
  summarise(barrels = sum(barrels),
            .groups = "drop") %>%
  mutate(total_barrels = cumsum(barrels)) %>%
  left_join(states, by = 'state')
```

Lets have a look at the materials trends. What ingredients are used in beer production?

```{r materialsTrends}
brewing_materials %>%
  filter(!str_detect(material_type, "Total"),
         !(month == 12 & year %in% 2014:2015)) %>%  # remove the subtotals) %>%
  filter(date == max(date)) %>%
  mutate(type = fct_reorder(type, month_current)) %>%
  ggplot(aes(type, month_current/1000000, fill =  type)) +
  geom_col() +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  coord_flip() +
  theme(legend.position = "",
        panel.grid.major.y =  element_blank()) +
  labs(
    x = '',
    y = 'Million Pounds',
    title = 'Ingredients in US Beer Production, in the month of December 2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )

```


```{r materialsTrendsChart}
brewing_materials %>%
  filter(!str_detect(type, "Total"), year < 2016,
         !(month == 12 & year %in% 2014:2015)) %>%  # remove the subtotals
  mutate(type = fct_reorder2(type, 
                             date,
                             month_current,
                             .fun = last2)) %>% 
  ggplot(aes(date, 
             month_current/1000000, 
             color = type)) +
  geom_line() +
    labs(
    x = '',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  ) +
  scale_radius(range = c(1,6)) +
  scale_x_date(date_breaks = "2 years",
               date_labels = "%Y") 

```

## Tidymetrics

```{r brewingSummarizedChart}

cross_by_dimensions <- function (tbl, ..., add = TRUE, max_dimensions = NULL, collect_fun = NULL) 
{
    g_vars <- dplyr::group_vars(tbl)
    columns <- ensyms(...)
    tbl <- tbl %>% ungroup() %>% mutate_at(vars(!!!columns), 
        as.character)
    if (!is.null(max_dimensions)) {
        tbl <- tbl %>% cross_by_dimensions_limited(
          columns,   
          max_dimensions = max_dimensions, 
          collect_fun = collect_fun)
    }
    else {
        for (column in columns) {
            tbl <- tbl %>% mutate(`:=`(!!column, "All")) %>% 
                union_all(tbl)
            if (!is.null(collect_fun)) {
                tbl <- collect_fun(tbl)
            }
        }
    }
    tbl %>% group_by_at(vars(all_of(g_vars))) %>% group_by(!!!columns, 
        .add = add)
}


brewing_summarized <- brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total"),
         year < 2016,
         !(month == 12 & year %in% 2014:2015)) %>%
  cross_by_dimensions(material_type, material) %>%  # duplicates the data to summarize it
  cross_by_periods(c("month", "quarter", "year")) %>%
  summarize(total_pounds = sum(month_current),
            .groups = "drop")

brewing_summarized %>%
  filter(material_type != "All", 
         material != "All", 
         period == "month") %>%
  mutate(material = fct_reorder2(material, 
                             date,
                             total_pounds,
                             .fun = last2)) %>% 
  ggplot(aes(date, 
             total_pounds/1000000, 
             color = material)) +
  geom_line() +
  scale_x_date(date_breaks = "2 years",
               date_labels = "%Y") +
  labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016:  Monthly',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
) 
```


```{r p4}
brewing_summarized %>%
  filter(material_type != "All", 
         material != "All", 
         period == "quarter") %>%
  mutate(material = fct_reorder2(material, 
                             date,
                             total_pounds,
                             .fun = last2)) %>% 
  ggplot(aes(date, 
             total_pounds/1000000, 
             color = material)) +
  geom_line() +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016: Quarterly',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  ) 
```

```{r p5}
brewing_summarized %>%
  filter(material_type != "All", 
         material != "All", 
         period == "year") %>%
  mutate(material = fct_reorder2(material, 
                             date,
                             total_pounds,
                             .fun = last2)) %>% 
  ggplot(aes(date, 
             total_pounds/1000000, 
             color = material)) +
  geom_line() +
    labs(
    x = 'Date',
    y = 'Million Pounds',
    color = "Material",
    title = 'Ingredients in US Beer Production, 2008-2016: Annually',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  )  
```

```{r materialsByYearChart}
brewing_materials %>%
  filter(material_type == "Total Used", year < 2016) %>%
  mutate(month = factor(month, labels = month.abb)) %>%
  ggplot(aes(x = month,
             y = month_current/1000000, 
             group = factor(year), 
             color = factor(year))) +
  geom_line() +
  expand_limits(y = 0) +
    labs(x = '',
    y = 'Million Pounds',
    color = "Year",
    title = 'Ingredients in US Beer Production, 2008-2016: Each Month',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday') +  
  scale_y_continuous(labels = scales::comma_format())  


```


```{r, shinymetrics, eval=FALSE}
remotes::install_github("ramnathv/shinymetrics")
remotes::install_github("ramnathv/shinybones")

brewing_metrics<-create_metrics(brewing_summarized)

library(shinymetrics)

# interactive plotly visualization of grouped time multi-dimensional BI 
preview_metric(brewing_metrics$production_NA_total_pounds)
# change the dropdown from "last year" to "all time" because 2019 is not in this dataset
```

```{r byBrewerSize}
brewer_size %>%
  filter(brewer_size != "Total", 
         !is.na(total_barrels)) %>%
  mutate(brewer_size = fct_lump_n(brewer_size, 
                                  n = 5, 
                                  w = total_barrels,
                                  other_level = "1 to 100,000 Barrels"),
         barrel_number = coalesce(parse_number(as.character(brewer_size)),1),
         brewer_size = fct_reorder(brewer_size, barrel_number)) %>% 
  ggplot(aes(as_factor(year), 
             total_barrels/1000000, 
             fill = brewer_size)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma_format())  +
    labs(x = 'Year',
    y = 'Million Barrels',
    fill = "Brewer Size",
    title = 'Total Production, by US Brewer Size in 2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday')  

```



## Where is Beer produced?

```{r whereBeerProduced}
beer_df %>% 
  ggplot(aes(year, barrels, 
                 group = state)) +
  geom_line(aes(color = state), show.legend = FALSE) +
  gghighlight(                               
    state %in% increase_states,
    label_key = state_name,
    label_params = list(size = 3, 
                        color = "black",
                        family = 'Arial Narrow',
                        fill = 'grey95', 
                        width = 2,
                        alpha = 1, 
                        label.r = 0.1, 
                        segment.alpha = 0),
    unhighlighted_params = list(color = 'grey88')
  ) +
  labs(
    x = 'Year',
    y = 'Total number of barrels produced',
    title = 'States with the Largest Increase in Beer Production, 2008-2019',
    subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
    caption = '@Jim_Gruman | #TidyTuesday'
  ) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_log10(labels = scales::label_number()) +
  theme(panel.grid = element_blank(),
        panel.grid.major = element_blank())

```

```{r BeerStatesMaps}
states_percents_2019 <- beer_states %>%
  filter( year == max(year), state != "total") %>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup() 

states <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))

states_percents_2019 %>%
   mutate(ID = if_else(state != "DC",
            str_to_lower(state.name[match(state,state.abb)]),
            "district of columbia")
          ) %>%
   inner_join(states, by = "ID") %>% 
   ggplot() +
   geom_sf(aes(geometry = geom, fill = percent)) +
   facet_wrap(~ type, nrow = 2) +
      scale_fill_continuous(labels = scales::percent) +
   labs(title = "In Each State, What is the Highest Proportion of Beer Consumed?",
        subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau',
        caption = '@Jim_Gruman | #TidyTuesday',
        fill = "% Percent") +
   theme_jim() +
   theme(legend.position = c(0.7,0.2),
         panel.grid.major = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank()) 

   
```

## An animated map.  Because it's TidyTuesday

```{r BeerStatesAnimated}
p <- beer_states %>%
  filter(state != "total") %>%
  group_by(state, year) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup() %>%
  mutate(ID = if_else(state != "DC",
            str_to_lower(state.name[match(state,state.abb)]),
            "district of columbia")
          ) %>%
  inner_join(states, by = "ID") %>% 
  ggplot(aes(geometry = geom, 
             fill = percent)) +
  geom_sf() +
  transition_time(year) +
  facet_wrap(~ type, nrow = 2) +
  labs(title = "In Each State, What is the Highest Proportion of Beer Consumed? {as.integer(frame_time)}",
        subtitle = 'US Alcohol and Tobacco Tax and Trade Bureau, 2008-2019',
        caption = '@Jim_Gruman | #TidyTuesday',
        fill = "% Percent") +
   theme_jim() +
   theme(legend.position = c(0.7,0.2),
         panel.grid.major = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank()) 

animate(p,  width = 900, height = 750, end_pause = 50, renderer = gifski_renderer())


```

## What are the relationships between the materials used? (using resampling methods)

For example, how much sugar do beer producers need per pound of malt?

```{r howMuchsugarPerMalt}
brewing_materials %>% 
  count(type, wt = month_current,  sort = TRUE) %>% 
    knitr::kable()

brewing_filtered <- brewing_materials %>%
  filter(!str_detect(type, "Total"), year < 2016,
         !(month == 12 & year %in% 2014:2015))

```

Explore the relationship between ingredients

```{r makeSugarMaltDF}
brewing_df <- brewing_filtered %>%
  select(date, type, month_current) %>%
  pivot_wider(names_from = type, 
              values_from = month_current) %>%
  janitor::clean_names()

# is there a remote correlation between
brewing_df %>%
  ggplot(aes(malt_and_malt_products, sugar_and_syrups)) +
  geom_smooth(method = "lm") +
  geom_point() +
  labs(title = "What is the relationship between malt and sugar ingredients?")
  
```


A simple linear model
```{r sugarMaltLinearModel}
# force the y-intercept to be zero (no beer, no malt, no sugar)
beer_fit <- lm(sugar_and_syrups ~ 0 + malt_and_malt_products, data = brewing_df)

summary(beer_fit)
```


```{r sugarMaltLinearModelTidyFits}
broom::tidy(beer_fit) %>% 
    knitr::kable()
```

## Bootstrap Resampling

to arrive at a confidence interval for a linear model

```{r setupBeerMaltBootstrap}
beer_boot <- bootstraps(data = brewing_df, times = 1e3, apparent = FALSE)
```

```{r beerMaltModelsOnBootstraps}
beer_models <- beer_boot %>% 
  dplyr::mutate(model = purrr::map(splits, ~ lm(sugar_and_syrups ~ 0 + malt_and_malt_products, 
                                  data = . )),
  coef_info = purrr::map(model, tidy))

beer_coefs <- beer_models %>%
     unnest(coef_info) 
     
```

## Evaluate results

What is the distribution of the estimates for the model?

```{r beerMaltSlopeEstimates}
beer_coefs %>%
  ggplot(aes(estimate)) +
  geom_histogram(alpha = 0.7) +
  labs(title = "Malt Coefficient Estimate for predicting Sugar and Syrups")
```

Our bootstrap confidence intervals:

```{r beerMaltSlopeConfidenceIntervals}
int_pctl(beer_models, coef_info)
```


```{r beerMaltSlopeCIplot}
beer_aug <- beer_models %>%
  dplyr::mutate(augmented = purrr::map(model, augment)) %>%
  unnest(augmented)

beer_aug %>%
  ggplot(aes(malt_and_malt_products, sugar_and_syrups)) +
  geom_line(aes(y = .fitted, group = id), alpha = 0.1, color = "cyan3") +
  geom_point() +
  labs(title = "Bootstrap Resampling Model Estimates Confidence Interval") 
```

And finally, my tidytuesday tweet:

```{r}
tweetrmd::include_tweet("https://twitter.com/jim_gruman/status/1246645041663221765")
```


