---
title: "Brain Injury"
author: "Jim Gruman"
date: 'March 24, 2020'
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
  html_document:
    toc: no
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE}
suppressPackageStartupMessages({
library(tidymodels)
library(tidyverse)
library(vip)
  
library(gganimate)
library(gifski)

})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))
```

## Get the Data

```{r loadData}
tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
tbi_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_year.csv')
tbi_military <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_military.csv')


```

# Data Exploration

Brain Injury Awareness Month, observed each March, was established 3 decades ago to educate the public about the incidence of brain injury and the needs of persons with brain injuries and their families (1). Caused by a bump, blow, or jolt to the head, or penetrating head injury, a traumatic brain injury (TBI) can lead to short- or long-term changes affecting thinking, sensation, language, or emotion.

The goal of this week's #TidyTuesday is to spread awareness for just how common TBIs are - both in civilian and military populations.

> One of every 60 people in the U.S. lives with a TBI related disability. Moderate and severe traumatic brain injury (TBI) can lead to a lifetime of physical, cognitive, emotional, and behavioral changes.


```{r incidenceChart}
tbi_year %>%
  ggplot(aes(factor(year), number_est, fill = factor(injury_mechanism))) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Brain Injury Incidence",
    subtitle = "US CDC",
    fill = "Injury",
    y = NULL,
    x = NULL,
    caption = paste0("Jim Gruman ", "#TidyTuesday March 24,2020")
  ) +
  theme(panel.grid.major.x = element_blank()) 
  
```


```{r BrainInjurygganimqdodge1}

p <- tbi_year %>%
  filter(injury_mechanism != "Total") %>%
  ggplot(aes(x = injury_mechanism, y = number_est, fill = type)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_x_discrete(
    labels = function(x)
      stringr::str_wrap(x, width = 20)
  ) +
  coord_flip() +
  transition_states(year, 
                    transition_length = 10, 
                    state_length = 1) +
  labs(
    title = "Traumatic Brain Injury (TBI) Incidence in {closest_state}",
    subtitle = "US CDC",
    x = "Cause of Injury",
    y = "Count",
    caption = paste0("Jim Gruman ", Sys.Date())
  ) +
  scale_y_log10(labels = scales::comma) 

animate(p,  
        width = 900, 
        height = 750, 
        end_pause = 50, 
        renderer = gifski_renderer())

```

