---
title: "Canadian Wind Turbines"
date: "October 27, 2020"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

The data this week comes from the [Government of Canada](https://open.canada.ca/data/en/dataset/79fdad93-9025-49ad-ba16-c26d718cc070).

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      dpi = 300, 
                      cache.lazy = FALSE,
                      tidy = "styler", 
                      fig.width = 12, 
                      fig.height = 7.402,
                      df_print = "paged")

library(tidyverse)
library(hrbrthemes)    # shortcuts for fonts and color palettes
extrafont::loadfonts()
library(tidytuesdayR) 

library(naniar)        # missing data exploration and imputation
library(visdat)        # handy missing data visuals

library(stringr)       # parse text nicely
library(kableExtra)    # features for nice table rendering
 
library(osmdata)       # Openstreetmap resources, like highways
library(sf)            # shapefile functions
library(rnaturalearth) # source of shapefile boundaries

library(ggtext)     # color annotations and plot labels
library(cowplot)   # combine plots into one graphic

theme_set(theme_ft_rc() +
            theme(plot.title.position = "plot"))

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-10-27")

```

```{r Readme, eval = interactive(), include=FALSE}

tt

```

## Glimpse the Data

Let's take an initial look at the format of the data available. There is a single table this week. Some records appear to be missing.

```{r Glimpse}

tt$`wind-turbine`

tt$`wind-turbine` %>% 
      vis_miss()
```

The turbine identifier and year of `objectid` 1451 is consistent with the other St. Lawrence `project_name` turbines. We will impute the missing value to the dataset as St. Lawrence.

Where turbine rated capacity is missing, the total rated capacity will be assumed to be spread uniformly across all turbines in a project if the model of the turbine is the same. This is the case for Armow Wind Project, South Kent Wind Farm, and Skyway8.

Commissioning_dates are spread in some cases across two or three or even as many as seven years. for the purposes of this analysis, I am going to parse out the right-most, last, date as the `allOnlineDate` as an integer.


```{r}
windTurbine <- tt$`wind-turbine` %>% 
   mutate(project_name = if_else(objectid == 1451,
                                 "St. Lawrence",
                                 project_name)) %>% 
  group_by(project_name) %>% 
  mutate(turbine_rated_capacity_k_w = 
              if_else(is.na(turbine_rated_capacity_k_w),
                      total_project_capacity_mw*1000/n(),
                      turbine_rated_capacity_k_w)) %>% 
  ungroup() %>% 
  mutate(allOnlineDate =  as.integer(str_sub(commissioning_date,
                          str_length(commissioning_date) - 3,
                          str_length(commissioning_date)))) %>% 
  mutate(manufacturer = if_else(str_detect(manufacturer, "Acciona"),
                                "Acciona",
                                manufacturer))

vis_dat(windTurbine)

```

# Visualize

The notes contain valuable insights. Some specifications may have changed since installation. In other cases, the dataset presented here has other technical assumptions.

```{r Visualize}
windTurbine %>% 
  filter(!is.na(notes)) %>% 
  distinct(project_name, notes, province_territory) %>% 
  knitr::kable(caption = "Canada Wind Farm Data Notes") %>% 
  kable_minimal(c("striped", "hover", "responsive"))
```

A count of turbines by manufacturer.

```{r manufacturers}

windTurbine %>% 
  group_by(manufacturer) %>% 
  summarize(Turbines = n(),
            `Total Installed Capacity in kW` = sum(turbine_rated_capacity_k_w)) %>% 
  arrange(desc(Turbines)) %>% 
  knitr::kable(caption = "Canada Wind Power Turbine Manufacturers") %>% 
  kable_minimal(c("striped", "hover", "responsive"))

```

Lets build an area chart showing the growth in capacity with an inset map:

```{r}
western_turbines <- filter(windTurbine, province_territory %in% c("Alberta", "British Columbia", "Manitoba", "Saskatchewan")) 

p1 <- western_turbines %>% 
  select(province_territory, allOnlineDate, turbine_rated_capacity_k_w) %>% 
   complete(allOnlineDate, province_territory, fill = list(turbine_rated_capacity_k_w = 0))  %>% 
  group_by(province_territory, allOnlineDate) %>% 
  summarize(sumCapacity = sum(turbine_rated_capacity_k_w)) %>% 
  mutate(CumSumCapacity = cumsum(sumCapacity)) %>%  
  ggplot() +
  geom_area(aes(allOnlineDate, CumSumCapacity/1000, fill = province_territory), show.legend = FALSE) +
  scale_y_continuous(labels = scales::comma_format(), 
                     position = "right",
                     expand = c(0,0)) +
  scale_x_continuous(limits = c(1995, 2019),
                     expand = c(0,0))+
  scale_fill_ipsum() +
  labs(y = "", x = "", fill = "",
  title = "Wind Power in Canada's Western Provinces",
  subtitle = "Cumulative sum of MW of power capacity installed at project completion",
  caption = "Data: Government of Canada | Visualization: @jim_gruman") 
  
```

Where are these turbines?

Can we take a look at turbines in the western provinces?

> adapted from Jake Kaupp's TidyTuesday submission this week on the Maritime provinces

```{r}
western_provinces <- ne_states("Canada", returnclass = "sf") %>% 
  filter(name %in% c("Alberta", "British Columbia", "Manitoba", "Saskatchewan"))

wp_box <- st_bbox(western_provinces) 

western_bbox <- matrix(c(wp_box[1], wp_box[3], wp_box[2], max(western_turbines$latitude) + 1), nrow = 2, ncol = 2)

western_hwy <- western_bbox %>% 
  opq(timeout = 6000) %>% 
  add_osm_feature(key = "highway", 
                  value = c("primary", "secondary", "tertiary", "secondary_link", "tertiary_link")) %>% 
  osmdata_sf()

western_hwy_clipped <- st_intersection(western_hwy$osm_lines, western_provinces)

p2 <- ggplot() +
  geom_sf(data = western_provinces, inherit.aes = FALSE, color = "#2F394D", fill = "white") +
  geom_sf(data = western_hwy_clipped, inherit.aes = FALSE, color = "#403d58", size = 0.1) +
  geom_point(data = western_turbines, aes(x = longitude, y = latitude, color = province_territory), alpha = 0.4, size = 4, show.legend = FALSE) +
  scale_color_ipsum() +
  geom_richtext(aes(x = -130, y = 58.5,
                    label = "The wind turbines built since 1995 as part of<br>larger wind farm projects in <span style='color:#8fd175'>British Columbia</span>,<br> <span style='color:#d18975'>Alberta</span>, <span style='color:#75b8d1'>Saskatchewan</span> and <span style='color:#3f2d54'>Manitoba</span>."), fill = "white", hjust = 0, label.color = NA, size = 5, color = "#464950", family = "Roboto Condensed") +
  labs(x = "", y = "") +
  scale_size_area(max_size = 6) +
  theme_map() 

```

```{r cowplot}
ggdraw(p1) +
  draw_plot(p2, 0, .3, .7, .7) 

```












