---
title: "Friends"
date: "September 8, 2020"
author: "Jim Gruman"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
  html_document:
    toc: no
    df_print: paged
editor_options:
  chunk_output_type: console
---

The data this week comes from the `friends` R package for the Friends transcripts and information on the episodes themselves, like when the characters interact with one another.

There's text, appearance, ratings, and many other datasets here. We will check out the `tidytext` mining package and the newly released [Supervised Machine Learning for Text Analysis in R](https://smltar.com/) book, both which are freely available online.

```{r setup, include=FALSE}

suppressPackageStartupMessages({
extrafont::loadfonts(device="win", quiet = TRUE)
library(tidyverse)
library(tidytext)
library(tidylo)
library(ragg)
library(showtext)
library(glue)
library(ggbump)
library(ggdist)
})

font_add(family = "Gabriel Weiss' Friends Font", 
         regular = "C:/Users/jimgr/AppData/Local/Microsoft/Windows/Fonts/GABRWFFR.TTF") 
font_add(family = "IBM Plex Sans", 
         regular = "C:/Users/jimgr/Documents/R/win-library/4.1/hrbrthemes/fonts/plex-sans/IBMPlexSans-Regular.TTF")

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))

friends_pal <- c("#F74035", "#008F48","#3F9DD4" , "#9787CD", "#F6D400", "#941205")

```

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tidytuesdayR::tt_load("2020-09-08")

```


```{r Wrangle}

friends_text <- tt$friends %>%
    inner_join(tt$friends_info, by = c("season", "episode")) %>%
    mutate(episode_title = glue("{ season }.{ episode } {title}"),
           episode_title = fct_reorder(episode_title, season + .001*episode),
           text = parse_character(text))

epi.info <- tt$friends_info
emotion <- tt$friends_emotions

main_characters <- friends_text %>%
  count(speaker, sort = TRUE) %>%
  head(6) %>%
  select(speaker) %>%
  separate(speaker, into = c("firstname", "lastname"), remove = FALSE)

main_characters %>%
  pull(speaker)

```

# Words each Character had per season

With inspiration from Dr. Christian Hoggard @CSHoggard

```{r parse}
friends_parsed <- friends_text %>% 
  filter(speaker %in% main_characters$speaker) %>%
  mutate(
    word_count = str_count(text, "\\w+"),
    speaker = word(speaker)
  ) %>%
  select(-utterance, -scene) %>% 
  group_by(speaker, season) %>%
  summarise(word_sum = sum(word_count),
            .groups = "drop")

friends_parsed %>% 
  group_by(speaker) %>%
  summarise(total_sum = sum(word_sum),
            .groups = "drop") %>% 
  knitr::kable()
```

```{r logo}
img_friends <- "https://turbologo.com/articles/wp-content/uploads/2019/12/friends-logo-cover-678x381.png."
```

```{r WhoSpokeTheMost, warning=FALSE}
p <-
  ggplot(friends_parsed, aes(season, word_sum, color = speaker)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  lims(y = c(5000, 14000)) +
  scale_x_continuous(n.breaks = 10) +
  scale_colour_manual(values = friends_pal) +
  labs(
    title = "Who Spoke the Most?",
    subtitle = "Number of words per season",
    caption = "@jim_gruman | Source: friends R package | #TidyTuesday",
    x = "Season",
    y = NULL
  ) +
  annotate(
    "text",
    x = 8,
    y = 13500,
    size = 4,
    color = "grey97",
    family = "IBM Plex Sans",
    label = "Rachel spoke over 17,000 words \n more than Pheobe!"
  ) +
  guides(colour = guide_legend(nrow = 1, 
                               override.aes = list(linetype = 0, size = 4))) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(color = "grey97", 
                               size = 11,
                               family = "IBM Plex Sans"),
    plot.margin = margin(20, 20, 20, 20),
    plot.background = element_rect(fill = "#000000"),
    panel.grid.major = element_line(
      size = 0.35,
      linetype = 'solid',
      colour = "grey10"
    ),
    panel.grid.minor = element_line(
      size = 0.3,
      linetype = 'solid',
      colour = "grey10"
    ),
    axis.text.x = element_text(color = "grey97",
                               family = "IBM Plex Sans"),
    axis.title.x = element_text(color = "grey97", 
                                margin = margin(20, 0, 5, 0),
                                family = "IBM Plex Sans"),
    axis.text.y = element_text(color = "grey97", 
                               margin = margin(0, 20, 0, 5),
                               family = "IBM Plex Sans"),
    axis.title.y = element_text(color = "grey97",
                                family = "IBM Plex Sans"),
    plot.title = element_text(
      color = "#9C61FD",
      size = 32,
      hjust = 0.5,
      margin = margin(0, 0, 10, 0),
      family = "Gabriel Weiss' Friends Font",
      face = "plain"
    ),
    plot.subtitle = element_text(
      color = "grey97",
      size = 12,
      hjust = 0.5,
      margin = margin(0, 0, 10, 0),
      family = "Gabriel Weiss' Friends Font",
      face = "plain"
    ),
    plot.caption = element_text(
      size = 9,
      colour = "grey97",
      margin = margin(20, 0, 0, 0),
      family = "IBM Plex Sans"
    )
  ) 

cowplot::ggdraw() +
  cowplot::draw_plot(p) +
  cowplot::draw_image(img_friends, scale = 0.2, y = -0.45, x = -0.4)


```

# Visualize the Lines by Season in Ranks

From David Smale @committedtotape

```{r}
tweetrmd::include_tweet("https://twitter.com/committedtotape/status/1303805992304627713")
```


```{r Ranks}
lines_by_season <- friends_text %>% 
  filter(speaker %in% main_characters$speaker) %>%
  count(season, speaker, name = "lines") %>%
  group_by(season) %>%
  arrange(season, -lines) %>%
  mutate(rank = row_number()) %>%
  inner_join(main_characters, by = "speaker") 


lines_by_season %>%
  ggplot(aes(season, rank, color = firstname)) +
  geom_point(size = 7) +
  geom_text(
    data = lines_by_season %>% filter(season == 1),
    aes(x = season - .2, label = firstname),
    size = 6,
    hjust = 1
  ) +
  geom_text(
    data = lines_by_season %>% filter(season == 10),
    aes(x = season + .2, label = firstname),
    size = 6,
    hjust = 0
  ) +
  geom_bump(size = 2, smooth = 5) +
  scale_x_continuous(
    limits = c(0, 11),
    breaks = seq(1, 10, 1),
    position = "top"
  ) +
  scale_y_reverse(breaks = seq(1, 6, 1),
                  position = "left") +
  scale_color_manual(values = friends_pal) +
  labs(
    title = "Aw, Phoebs!",
    subtitle = "Friends ranked by number of lines per season \nPhoebe has the fewest lines in 5 of 10 seasons",
    caption = "Data: friends R package ",
    x = "Season",
    y = NULL
  ) +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "#000000", color = "gray10"),
    axis.text.x = element_text(color = "gray90", size = 16),
    axis.text.y = element_text(color = "gray90", size = 16),
    axis.title.x.top = element_text(
      color = "gray90",
      size = 16,
      hjust = 0.5,
      margin = margin(0, 0, 10, 0)
    ),
    axis.title.y.left = element_text(
      color = "gray90",
      size = 16,
      hjust = 0,
      vjust = 0.5,
      angle = 0
    ),
    plot.title = element_text(
      color = "#9C61FD",
      hjust = 0.5,
      size = 34,
      margin = margin(0, 0, 10, 0),
      family = "Gabriel Weiss' Friends Font",
      face = "plain"
    ),
    plot.subtitle = element_text(
      color = "gray90",
      hjust = 0.5,
      size = 16,
      margin = margin(0, 0, 20, 0),
      family = "Gabriel Weiss' Friends Font",
      face = "plain"
    ),
    plot.caption = element_text(
      color = "gray90",
      hjust = 1,
      size = 10,
      margin = margin(20, 0, 5, 0)
    ),
    plot.margin = margin(20, 20, 20, 20),
    panel.grid.major = element_line(
      size = 0.35,
      linetype = 'solid',
      colour = "grey10"
    ),
    panel.grid.minor = element_line(
      size = 0.3,
      linetype = 'solid',
      colour = "grey10"
    )
  )

```

Let's model features with regression to infer what features drive imdb_rating

```{r LinesPerEpisode}

speaker_lines_per_episode <- friends_text %>% 
  count(speaker, episode_title, imdb_rating) %>% 
  complete(speaker, episode_title, fill = list(n = 0)) %>% 
  group_by(episode_title) %>%
  fill(imdb_rating, .direction = "downup") %>% 
  ungroup() %>% 
  add_count(episode_title, wt = n, name = "episode_total") %>% 
  mutate(pct = n / episode_total)

speaker_lines_per_episode %>%
  semi_join(main_characters, by = "speaker") %>%
  mutate(speaker = fct_reorder(speaker, n)) %>% 
  ggplot(aes(pct, 
             speaker,
             color = speaker,
             fill = speaker)) +
  ggdist::stat_dots(show.legend = FALSE,
                    side = "both",
                    layout = "weave") +
  ggrepel::geom_text_repel(data = . %>% filter(pct > 0.32),
            aes(label = episode_title),
            direction = "x",
            show.legend = FALSE) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "", y = "",
       title = "Lines portion within each episode",
       fill = function(x) paste0(as.numeric(x)*100, "%"),
       caption = "Data: friends R package | Visual by @jim_gruman") +
  coord_flip() +
  theme(
      plot.title = element_text(
      color = "#9C61FD",
      hjust = 0.5,
      size = 34,
      margin = margin(0, 0, 10, 0),
      family = "Gabriel Weiss' Friends Font",
      face = "plain"
    ),
    plot.background = element_rect(fill = "#000000", color = "gray10"),
    plot.caption = element_text(
      color = "gray90",
      hjust = 1,
      size = 10,
      margin = margin(20, 0, 5, 0)
    ),
    plot.margin = margin(20, 20, 20, 20),
    panel.grid.major = element_line(
      size = 0.35,
      linetype = 'solid',
      colour = "grey10"
    ),
    panel.grid.minor = element_line(
      size = 0.3,
      linetype = 'solid',
      colour = "grey10"
    )
)

speaker_lines_per_episode %>% 
  semi_join(main_characters, by = "speaker") %>% 
  group_by(speaker) %>% 
  summarize(correlation = cor(pct,imdb_rating)) %>% 
  knitr::kable()

# setup the training set
speakers_per_episode_wide <- speaker_lines_per_episode %>% 
  semi_join(main_characters, by = "speaker") %>% 
  select(episode_title, speaker, pct, imdb_rating, episode_total) %>% 
  spread(speaker, pct) %>%
  select(-episode_title)

speakers_per_episode_wide %>% 
  lm(imdb_rating ~ ., data = .) %>% 
  summary() 

```

The more Ross speaks, the more popular the episode is

Let's look at words, and weighted log-odds ratios

```{r uniqueSentiments}

words_unnested <- friends_text %>%
   left_join(emotion, by = c("season", "episode", "utterance")) %>% 
   select(text, speaker, season, episode_title, emotion) %>% 
   unnest_tokens(word, text, to_lower = FALSE) %>% 
   anti_join(stop_words, by = "word") %>% 
   filter(!word %in% c("yeah","hey","gonna","uh", "y'know", "um", "ah", "um", "la", "wh"),
          !str_detect(word, "[:digit:]|^aa"),
          !is.na(emotion), 
          emotion != "Neutral" ) %>% 
  mutate(word = str_to_lower(word))

words_unnested %>% 
  count(word, sort = TRUE) %>% 
  slice_max(order_by = n, n = 15) %>% 
  knitr::kable()

by_speaker_word <- words_unnested %>%
  semi_join(main_characters, by = "speaker") %>% 
  count(speaker, word, emotion, name = "n_emotion") %>% 
  group_by(speaker, word) %>% 
  summarize(n = sum(n_emotion),
            emotion =  first(emotion, order_by = -n_emotion),
            .groups = "drop") 

# what are the most over-represented words for each character?
unique_sentiments <-  by_speaker_word %>% 
  bind_log_odds(speaker, word, n) %>% 
  group_by(speaker) %>% 
  slice_max(log_odds_weighted, n = 12) %>% 
  mutate(size = scales::rescale(log_odds_weighted, to = c(3,9)))

```

# Visualize a word cloud with words most representative of each character and most common emotional sentiment

Jack Davison @JDavison_ inspired this beautiful word cloud

Credit David Robinson for the `tidylo` insight 

```{r word_cloud, fig.asp = 0.7}
ggplot(unique_sentiments,
       aes(
         label = word,
         size = log_odds_weighted,
         color = emotion,
         alpha = log_odds_weighted
       )) +
  ggwordcloud::geom_text_wordcloud_area(area_corr_power = 1) +
  facet_wrap(~ speaker) +
  scale_radius(range = c(3, 20)) +
  scale_color_manual(values = friends_pal) +
  scale_alpha(range = c(.5, 1)) +
  theme(
    plot.background = element_rect(fill = "#000000", color = NA),
    strip.text = element_text(
      family = "Gabriel Weiss' Friends Font",
      size = 20,
      color = "white",
      hjust = 0.5
    ),
    plot.margin = unit(rep(1, 4), "cm"),
    panel.spacing = unit(.5, "cm"),
    plot.title = element_text(
      family = "Gabriel Weiss' Friends Font",
      face = "plain",
      size = 32,
      color = "#f4c93cff",
      hjust = .5,
      vjust = .5
    ),
    plot.subtitle = ggtext::element_markdown(
      hjust = .5,
      color = "white",
      size = 15
    ),
    plot.caption = ggtext::element_markdown(
      hjust = .5,
      vjust = .5,
      color = "#f4c93cff"
    )
  ) +
  labs(title = "the one with the sentiment analysis",
       subtitle = "<span style='color:#F74035'>Joyful</span> <span style='color:#008F48'>Mad</span> <span style='color:#3F9DD4'>Peaceful</span> <span style='color:#9787CD'>Powerful</span> <span style='color:#F6D400'>Sad</span> <span style='color:#941205'>Scared</span><br><br>",
       caption = "<br>Data from <b>{friends}</b> (github.com/EmilHvitfeldt/friends)<br> Visualisation by <b>Jim Gruman</b> (Twitter @jim_gruman)<br>Code found at <b>opus1993.github.io/myTidyTuesday</b>")


```

My #TidyTuesday tweet:

```{r}
tweetrmd::include_tweet("https://twitter.com/jim_gruman/status/1304230550081875968")
```






