---
title: "Global Transit Costs"
author: "Jim Gruman"
date: "January 6, 2021"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

Our goal here is to explore data from the [Transit Costs Project](https://transitcosts.com/). The organization has created a database that spans more than 50 countries and totals more than 11,000 km of urban rail built since the late 1990s. 

Why do transit-infrastructure projects in New York cost 20 times more on a per kilometer basis than in Seoul? Let's look for insights into what the cost drivers are.

```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 7.402,
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	cache.lazy = FALSE,
	df_print = "paged",
	dpi = 300,
	tidy = "styler"
)
```

```{r load-packages}
suppressPackageStartupMessages({
library(tidyverse)
library(lubridate)
library(here)
library(WDI)
library(tidymodels)
library(baguette)
library(themis)
})
```

```{r load-fonts}
extrafont::loadfonts(quiet = TRUE)

theme_set(hrbrthemes::theme_ipsum_gs() +
            theme(plot.title.position = "plot", 
                  plot.caption.position = "plot"))
```

We will load the latest data directly from the #TidyTuesday github repo.

```{r load_data}
raw  <- tidytuesdayR::tt_load("2021-01-05")

transit_cost <- raw$transit_cost
```

Let's also bring in a recent GDP per capita figure:

```{r}
missing <- tibble(country = "TW",
       income = "High income",
       gdp_per_capita = 54020,
       pop = 23340000)

country_incomes <- WDI(
  indicator = c(
    gdp_per_capita = "NY.GDP.PCAP.PP.KD",
    pop = "SP.POP.TOTL"
  ),
  start = 2019,
  end = 2019,
  extra = TRUE
) %>% 
  as_tibble() %>%
  select(country = iso2c, income, gdp_per_capita, pop) %>%
  filter(!is.na(income)) %>%
  bind_rows(missing) %>%  
  mutate(income = as_factor(income)) %>%
  mutate(country = if_else(country == "GB","UK",country)) %>% 
  mutate(income = fct_relevel(
    income,
    "Aggregates",
    "Low income",
    "Lower middle income",
    "Upper middle income"
  )) 

```

Prepare the data 

```{r}
transit_df <- transit_cost %>% 
  filter(!is.na(real_cost), !is.na(year), end_year != "X", real_cost > 0) %>% 
  left_join(country_incomes, by = "country" ) %>% 
  mutate(build_year = case_when(
            start_year == "5 years" ~ "2021",
            start_year == "4 years" ~ "2021",
            start_year == "not start" ~ "2022",
            !is.na(start_year) ~ start_year,
            !is.na(year) ~ as.character(year),
            TRUE ~ "Look"),
         build_year = as.integer(build_year),
         build_time = as.integer(end_year) - build_year,
         length = if_else(is.na(length),0,length),
         tunnel = if_else(is.na(tunnel),0,tunnel),
         nonTunnelLength = length - tunnel,
         build_year = as.integer(build_year),
         stations = if_else(is.na(stations),0,stations),
         rr = if_else(is.na(rr),0,rr),
         regional = factor(rr, labels = c("urban", "intracity")),
         location = fct_lump(glue::glue("{city}, {country}"), 25),
         cost = as.numeric(real_cost)) %>% 
  rename(TunnelLength = tunnel) %>% 
  select(cost, build_year, location, regional, nonTunnelLength, TunnelLength, stations, income, build_time) 


```

Let's look at the distributions of the independent variables:

```{r}
transit_df %>% 
  select(-cost) %>% 
  keep(is.numeric) %>%
  gather() %>%
  ggplot() +
  geom_histogram(mapping = aes(x = value, fill = key), color = "black") +
  facet_wrap(~ key, scales = "free") +
  scale_x_continuous(n.breaks = 3) +
  theme(legend.position = "",
        plot.title.position = "plot") +
  labs(title = "Global Transit Costs Numeric Variable Histogram Distributions")
```

The lengths and stations both skew right. What happens if we look at the same with an log function applied to x?

```{r}
transit_df %>% 
  select(-cost, -build_year) %>% 
  keep(is.numeric) %>%
  gather() %>%
  ggplot() +
  geom_histogram(mapping = aes(x = value, fill = key), color = "black", show.legend = FALSE) +
  scale_x_continuous(n.breaks = 3, trans = "log2") +
  facet_wrap(~ key, scales = "free") +
  labs(title = "Global Transit Costs Numeric Variable Histogram Distributions")
```

Another data visualization showing the relationships between the numeric features and cost:

```{r}
transit_df %>%
  keep(is.numeric) %>%
  pivot_longer(-cost, names_to = "Feature", values_to = "Value") %>%
  ggplot() +
  geom_point(aes(x = Value, y = cost, color = Feature), show.legend = FALSE) + 
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~ Feature, scales = "free") +
  scale_x_continuous(n.breaks = 3, trans = "log2") +
  labs(x = "Numeric Feature Value",
       title = "Global Transit Costs Numeric Variable versus Cost")
```

Applying log transformations to length, station, and tunnel could be helpful.

What are the numeric variable correlations?

```{r}
transit_df %>%
  keep(is.numeric) %>%
  corrr::correlate() %>%
  corrr::network_plot(min_cor = 0.2)
```

There is a risk that there will be hidden interactions between numeric features.

Let's explore the categorical variables.

```{r}
transit_factors <- transit_df %>% keep(is.factor) %>% colnames()

chart <- c(transit_factors, "cost")

transit_df %>%
  select_at(vars(chart)) %>%
  pivot_longer(-cost, names_to = "Factor", values_to = "Level") %>%
  mutate(Factor = fct_reorder(Factor, cost),
         Level = fct_reorder(Level, cost)) %>% 
  ggplot() +
  geom_boxplot(aes(x = Level, y = cost, fill = Factor), show.legend = FALSE) + 
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  facet_wrap(~ Factor, scales = "free") +
  labs(x = "Categorical Feature Value",
       title = "Global Transit Categorical Variable versus Cost")

```

Paris, France and Doha (the costly "Other") may be cost outliers. I wonder if the numeric variables will fully explain the true cost. Out of curiosity, what is the dataset implying about the length of a regional train?

```{r}
transit_df %>% 
  ggplot() +
  geom_boxplot(aes(y = regional, 
                   x = TunnelLength + nonTunnelLength, 
                   fill = regional), show.legend = FALSE) +
  scale_x_continuous(trans = "log2", labels = scales::comma_format()) +
  labs(title = "Regional train projects are not necessarily of longer length.")

```


## Setup for modeling

split the data into training and testing sets

```{r}
transit_split <- initial_split(transit_df)

transit_split
```

```{r}
transit_train <- training(transit_split)
transit_test <- testing(transit_split)
```

At some point we’re going to want to do some parameter tuning, so we’re going to use cross-validation. We can create a cross-validated version of the training set in preparation for that moment using `vfold_cv()`.

```{r}
transit_folds  <- vfold_cv(transit_train, v = 10, strata = "cost")
```

Let's craft a recipe

```{r}
transit_rec <- recipe(cost ~ ., data = transit_train) %>% 
  step_normalize(all_numeric()) %>% 
  step_BoxCox(nonTunnelLength,stations,TunnelLength) %>% 
  step_dummy(all_nominal()) %>% 
  step_interact( ~ stations:nonTunnelLength)

transit_prep <- prep(transit_rec)

```

## Specify a model

```{r}
linear_spec <- linear_reg() %>%
  set_engine("glmnet") %>%
  set_mode("regression")

transit_wflow <- workflow() %>%
  add_model(linear_spec) %>% 
  add_recipe(transit_rec)
  
transit_wflow
```


```{r eval = FALSE}
doParallel::registerDoParallel()

crash_res <- fit_resamples(
  transit_wflow,
  transit_folds
)

```

