---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
    code_folding: hide
editor_options:
  chunk_output_type: console
---

Welcome to my TidyTuesday website collection. Most of the activity here is related to the [rfordatascience #TidyTuesday](https://github.com/rfordatascience/tidytuesday) weekly social data project. The exercises here are certainly not finished products. They demonstrate what could be accomplished in an hour or two after-hours from datasets that can be found out in the "wild." In many cases, I've used these to explore packages with new features and functions and to facilitate conversations with others in the community via Twitter and Slack.

```{r, include = FALSE}
library(purrr)
library(dplyr)
library(knitr)
library(kableExtra)
library(stringr)

knitr::opts_chunk$set(cache = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      dpi = 300, 
                      cache.lazy = FALSE,
                      tidy = "styler", 
                      fig.width = 12, 
                      fig.height = 7.402)

articles <- tibble(file = dir(path = "analysis")) %>% 
  filter(str_detect(file, "\\.Rmd")) %>% 
  filter(!file %in% c("_site.yml","about.Rmd","index.Rmd","license.Rmd","rsconnect"))

title_f <- function(file) {
  lines <- readLines(paste0("analysis/",file),
                     warn = FALSE)  
  lines[grepl("^title:", lines)]
}

date_f <- function(file) {
  lines <- readLines(paste0("analysis/",file),
                     warn = FALSE)  
  lines[grepl("^date:", lines)]
}

library_f <- function(file) {
  lines <- readLines(paste0("analysis/",file),
                     warn = FALSE)  
  lines[grepl("^library", lines)]
}
```

```{r}

articlesTable <- articles %>% 
  mutate(title = map_chr(file,~ title_f(.))) %>% 
  mutate(dateline = map_chr(file, ~ date_f(.))) %>% 
  mutate(libraries = map(file, ~ library_f(.))) %>%
  tidyr::unnest_longer(col = libraries) %>% 
  mutate(title = str_sub(title, 
                         start = 9L,
                        end = str_length(title) - 1L)) %>% 
  mutate(dateline = str_sub(dateline, 
                            start = 8L,
                            end = str_length(dateline) - 1L)) %>% 
  mutate(libraries = str_squish(str_sub(libraries, 
                             start = str_locate(libraries, "\\(")[,1] + 1,
                             end = str_locate(libraries, "\\)")[,1] - 1 ))) %>% 
  filter(!libraries %in% c("tidyverse","tidytuesdayR","lubridate","here","hrbrthemes", "tweetrmd", "grumanlib")) %>% 
  arrange(desc(lubridate::mdy(dateline)),
          libraries) %>% 
  tidyr::nest(package = libraries) %>% 
  mutate(link = stringr::str_remove(file,".Rmd")) %>% 
  mutate(link = paste0("https://opus1993.github.io/myTidyTuesday/",link,".html"))

articlesTable %>% 
  mutate(title = text_spec(title,
                           format = "html", 
                           link = link)) %>% 
  select(-file, -link) %>% 
  knitr::kable("html", 
               escape = FALSE,
               col.names = c("Title","Date","Packages")) %>%
  kable_styling(bootstrap_options = c("hover", "condensed"),
                full_width = TRUE,
                fixed_thead = TRUE) %>% 
  column_spec(1:3,width = c("10em","10em","60em"))

```

----

I have made notes of my sources for code and content. Please be sure to reach out, acknowledge, and cite their contributions directly.
