---
title: "Historical Phone Usage"
date: "November 10, 2020"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

The data this week comes from  [OurWorldInData.org](https://ourworldindata.org/technology-adoption) and is related to telephone subscriptions, mobile and landline, as part of the **Adoption of technologies around the world** series.

Their web page presents high quality interactive javascript visuals of time series line charts and map chloropleths as

<iframe src="https://ourworldindata.org/grapher/fixed-telephone-subscriptions-per-100-people?tab=chart" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>

```{r include = FALSE}
htmltools::includeHTML("https://ourworldindata.org/grapher/fixed-telephone-subscriptions-per-100-people?tab=chart")
```

and

<iframe src="https://ourworldindata.org/grapher/mobile-cellular-subscriptions-per-100-people?tab=chart" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>

```{r include = FALSE}
htmltools::includeHTML("https://ourworldindata.org/grapher/mobile-cellular-subscriptions-per-100-people?tab=chart")
```

Our mission is to discover other insights with the dataset

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)    # shortcuts for fonts and color palettes
extrafont::loadfonts(quiet = TRUE)
library(tidytuesdayR) 

library(scales)

library(WDI)          # World Bank World Development Indicators

library(naniar)        # missing data exploration and imputation
library(visdat)        # handy missing data visuals

library(ggtext)     # color annotations and plot labels

library(ggrepel)
})
source(here::here("code","_common.R"),  
       verbose = FALSE,
       local = knitr::knit_global())
ggplot2::theme_set(theme_jim(base_size = 12))

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-11-10")

```

```{r Readme, eval = interactive(), include=FALSE}

tt

```

Let's take an initial look at the format of the data available. There are two tables included.  Some of the gdp data is missing

The subs figure is a rate of subscriptions per 100 population

Mobile data is only available to 2017. To make meaningful comparisons, we will stack the datasets and trim to 2017

```{r}
phones <- bind_rows(tt$mobile %>% 
  rename(subscriptions = mobile_subs) %>% 
  mutate(type = "Mobile") , 
  tt$landline %>% 
  rename(subscriptions = landline_subs) %>% 
  mutate(type = "Landline")) %>% 
  rename(country = entity) %>% 
  mutate(type = fct_relevel(type, c("Mobile","Landline")),
         year = as.integer(year)) %>% 
  filter(year <= 2017L, !str_detect(country, "Serbia")) %>% 
  complete(type, 
           nesting(continent, country), 
           year = 1990L:2017L) %>% 
  group_by(country) %>% 
  tidyr::fill(subscriptions) %>% 
  ungroup() %>% 
  tidyr::replace_na(list(subscriptions = 0))
```

For the United States:

```{r}
phones %>% 
  filter(country == "United States", !is.na(subscriptions)) %>% 
  ggplot(aes(year, subscriptions, color = type)) +
  geom_line(size = 2) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  scale_linetype_manual(guide = guide_legend(reverse = FALSE)) +
  labs(title = "Telephone subscriptions in the United States",
      subtitle = "Rate / 100 Population", x = "", y = "", color = "") +
    theme(legend.position = c(0.9,0.5),
        legend.background = element_rect(color = "white"),
        axis.line.x = element_blank()) 

```

For the single nation of Sudan:

```{r}
phones %>% 
  filter(country == "Sudan", !is.na(subscriptions)) %>% 
  ggplot(aes(year, subscriptions, color = type)) +
  geom_line(size = 2) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  scale_linetype_manual(guide = guide_legend(reverse = FALSE)) +
  labs(title = "Telephone subscriptions in Sudan",
      subtitle = "Rate / 100 Population", x = "", y = "", color = "") +
  annotate("text", x = 2002, y = 20, 
           label = "Sudan skipped landline \ninfrastructure development",
           size = 4,
           color  = "black") +
    theme(legend.position = c(0.9,0.3),
        legend.background = element_rect(color = "white"),
        axis.line.x = element_blank()) 
```

For the single nation of Luxembourg:

```{r}
phones %>% 
  filter(country == "Luxembourg", !is.na(subscriptions)) %>% 
  ggplot(aes(year, subscriptions, color = type)) +
  geom_line(size = 2) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  scale_linetype_manual(guide = guide_legend(reverse = FALSE)) +
  labs(title = "Telephone subscriptions in Luxembourg",
      subtitle = "Rate / 100 Population", x = "", y = "", color = "") +
  annotate("text", x = 1995, y = 43, 
           label = "Around 50 of 100 people \n retain landlines",
           size = 4,
           color  = "black") +
  annotate("text", x = 2010, y = 120, 
           label = "There are more moble subscriptions \nthan there are people",
           size = 4,
           color  = "black") +
      theme(legend.position = c(0.9,0.5),
        legend.background = element_rect(color = "white"),
        axis.line.x = element_blank()) 

```

For the single nation of North Korea:

```{r}
phones %>% 
  filter(country == "North Korea", !is.na(subscriptions)) %>% 
  ggplot(aes(year, subscriptions, color = type)) +
  geom_line(size = 2) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  scale_linetype_manual(guide = guide_legend(reverse = FALSE)) +
  annotate("text", x = 2010, y = 11, 
           label = "2017 is unknown",
           size = 4,
           color  = "black") +
  labs(title = "Telephone Subscribers in North Korea",
      subtitle = "Rate / 100 Population", x = "", y = "", color = "") +
      theme(legend.position = c(0.9,0.5),
        legend.background = element_rect(color = "white"),
        axis.line.x = element_blank()) 

```

Descriptive statistics across continents and by national income:

```{r}

summarize_subscriptions <- . %>%        # making the summary re-usable
  filter(!is.na(subscriptions)) %>%
  summarize(avg_subscriptions = mean(subscriptions),
            median_subscriptions = median(subscriptions),
            q25 = quantile(subscriptions, .25),
            q75 = quantile(subscriptions, .75), 
            .groups = "drop")

phones %>%
  filter(!is.na(subscriptions)) %>%
  group_by(year, continent, type) %>%
  summarize_subscriptions() %>%
  ggplot(aes(year,
             avg_subscriptions,
             color = type)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = .2, show.legend = FALSE) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  facet_wrap(~ continent) +
  theme(legend.position = c(0.8, 0.2)) +
  labs(x = "",
       y = "Median subscriptions per person",
       color = "",
       title = "How do mobile and landline adoption differ between continents?",
       subtitle = "Ribbon shows the 25th-75th percentile range of all countries",
      caption = "Data: OurWorldInData.org | Visual: @jim_gruman")
```


```{r}
country_incomes <- WDI(indicator = c(gdp_per_capita = "NY.GDP.PCAP.PP.KD",
                                     pop = "SP.POP.TOTL"),
                       start = 2005, 
                       end = 2005, 
                       extra = TRUE) %>%
  as_tibble() %>%
  select(code = iso3c, income, gdp_per_capita, pop) %>%
  filter(!is.na(income)) %>% 
  mutate(income = as_factor(income))  %>%
  mutate(income = fct_relevel(income, 
                              "Aggregates",
                              "Low income", 
                              "Lower middle income", 
                              "Upper middle income"))

by_year_income <- phones %>%
  inner_join(country_incomes, by = "code") %>%
  group_by(year, income, type) %>%
  summarize_subscriptions()

by_year_income %>%
  ggplot(aes(year,
             median_subscriptions,
             color = type)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = q25, ymax = q75), 
              alpha = .2, show.legend = FALSE) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  facet_wrap(~ income) +
  labs(
       y = "Median subscriptions per person", x = "",
       color = "",
       title = "How do mobile and landline adoption differ between national income categories?",
       subtitle = "Ribbon shows the 25th-75th percentile range of all countries",
      caption = "Data: OurWorldInData.org and World Bank | Visual: @jim_gruman") +
  theme(legend.position = "bottom")

```


What are ways we could summarize stats for each country?

```{r}
country_sizes <- phones %>%
  group_by(country) %>%
  summarize(avg_population = mean(total_pop, na.rm = TRUE),
            .groups = "drop" ) %>% 
  slice_max(order_by = avg_population, n = 52) 

phones %>%
  semi_join(country_sizes, by = "country") %>% 
  ggplot(aes(year,
             subscriptions,
             color = type,
             group = interaction(type, country))) +
  geom_line(size = 1.5) +
  scale_x_continuous(expand = expansion(mult = c(0,0)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = NULL) +
  facet_wrap(~ continent) +
  geom_hline(yintercept = 70, lty = 2) +
  labs(y = "Subscriptions per 100 population",
       x = "",
       title = "Mobile phone technology penetration and the race to 70 subscriptions per 100 people",
       subtitle = "In the largest 50 nations, the US, Canada, and Europe replaced landlines. Africa started from scratch with mobile",
       color = "",
      caption = "Data: OurWorldInData.org and World Bank | Visual: @jim_gruman") +
  theme(legend.position = c(0.9, 0.2))

```

```{r}
countries_summarized <- phones %>%
  filter(!is.na(subscriptions)) %>%
  select(-total_pop, gdp_per_cap) %>%
  pivot_wider(names_from = type, values_from = subscriptions) %>%
  group_by(continent, country, code) %>%
  summarize(year_past_70_mobile = na_if(min(year[Mobile >= 70]), Inf),
            peak_landline = max(Landline, na.rm = TRUE),
            peak_mobile = max(Mobile),
            n_mobile = sum(!is.na(Mobile))) %>%
  ungroup() %>%
  inner_join(country_incomes, by = "code") %>%
  filter(n_mobile >= 25) %>%
  arrange(desc(year_past_70_mobile))

countries_summarized %>%
  tidyr::drop_na() %>% 
  ggplot(aes(year_past_70_mobile,
             gdp_per_capita,
             color = continent,
             size = pop)) +
  geom_point(alpha = 0.6) +
  geom_text_repel(aes(label = country), show.legend = FALSE) +
  scale_y_log10(labels = dollar) +
  scale_x_continuous(expand = expansion(mult = c(0.05,0.05)),
    breaks = seq(1990, 2020, 5),
    minor_breaks = seq(1990, 2019, 1)) +
  scale_color_discrete(guide = guide_legend(override.aes = list(size = 10))) +
  scale_size_continuous(labels = comma,
                        range = c(3,8)) +
  labs(y = "GDP per capita (in 2005)",
       x = "Year mobile subscriptions passed 70/100 people",
       title = "Wealthy countries in Europe had early mobile phone technology penetration",
       subtitle = "The year that each nation first exceeded 70 subscribers per 100 population",
       size = "National \nPopulation",
       color = "",
      caption = "Data: OurWorldInData.org and World Bank | Visual: @jim_gruman") +
  theme(legend.position = "bottom",
        legend.background = element_rect(color = "white")) 

```






