---
title: "US Poverty"
date: "November 11, 2020"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

The data here comes from the [US Census American Community Survey](https://www.census.gov/library/visualizations/interactive/2014-2018-poverty-rate-by-county.html) 2014-2018 5-year dataset.

Our goal is to recreate this beautiful Tableau [graphic](https://public.tableau.com/views/18_5yr_Poverty/2018Poverty?:embed=y&:embed_code_version=3&:loadOrderID=0&:display_count=y&:origin=viz_share_link), but to represent the population size better.

<iframe src="https://public.tableau.com/views/18_5yr_Poverty/2018Poverty?:language=en&:display_count=y&:origin=viz_share_link" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>


```{r setup}

suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)    # shortcuts for fonts and color palettes
extrafont::loadfonts(quiet = TRUE)

library(cartogram)
library(sf)
library(tmap)
library(tigris)

library(tidycensus)
})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())

ggplot2::theme_set(theme_jim(base_size = 12))

```


```{r getData, message=FALSE, warning=FALSE}
nonx <- c('78', '69', '66', '72', '60', '15', '02')

if (file.exists("data/us_states.rds")) {
  us_states <- read_rds("data/us_states.rds")
} else {
  us_states <- states(cb = TRUE) %>%
    filter(!STATEFP %in% nonx) %>%
    st_transform(crs = 2163)
  
  write_rds(us_states, "data/us_states.rds")
}


if (file.exists("data/acs_poverty.rds")) {
  poverty <- read_rds("data/acs_poverty.rds")
} else {
  poverty <- tidycensus::get_acs(
    geography = "county",
    variables = c(poverty = "B17013_001"),
    geometry = TRUE,
    year = 2018
  ) %>%
    select(GEOID, NAME, estimate, moe, geometry)
  
  write_rds(poverty, "data/acs_poverty.rds")
}

population <-
  suppressMessages(
    tidycensus::get_acs(
      geography = "county",
      variables = c(population = "B01003_001"),
      geometry = FALSE,
      year = 2018
    ) %>%
      select(GEOID, estimate)
  )
```

The bubbles are generated on top of the shapefile centroids, and then adjusted to fit through iterations with the cartogram package.

```{r}
counties_sf <- poverty %>% 
  left_join(population, by = "GEOID", suffix = c(".pov",".pop")) %>% 
  mutate(rate = round(estimate.pov/estimate.pop,2),
         moe_rate = round(moe/estimate.pop,3)) %>% 
  st_transform(crs = 2163)

county_dorling <-
  cartogram_dorling(
    x = counties_sf,
    weight = "estimate.pop",
    k = 0.2,
    itermax = 100
  )
```

## A ggplot

```{r ggplot}
county_dorling %>% 
  filter(!str_detect(NAME, "Alaska|Hawaii|Puerto|Guam")) %>% 
  ggplot(aes(fill = rate)) +
  geom_sf(color = "grey60") +
  geom_sf(data = us_states, 
          fill = NA, 
          show.legend = F, 
          color = "grey60", 
          lwd = .5 ) +
  coord_sf(crs = 2163, datum = NA) +
  scale_fill_stepsn(colors = c('#eff3ff','#bdd7e7','#6baed6','#3182bd','#08519c')) +
  labs(title = "2014-2018 Poverty Rate in the United States by County", subtitle = "Bubble size corresponds to County Population",
       caption = "US Census American Community Survey | jim_gruman")


```

And the original tweet, as posted to twitter:

```{r}
tweetrmd::include_tweet("https://twitter.com/jim_gruman/status/1326673815288623105")

```
