---
title: "Washington State Trails"
author: "Jim Gruman"
date: "November 27, 2020"
output:
  workflowr::wflow_html:
    toc: no
    code_folding: hide
editor_options:
  chunk_output_type: console
---

The data this week comes from [Washington Trails Association](https://www.wta.org/go-outside/hikes?b_start:int=1) courtesy of the [TidyX crew](https://github.com/thebioengineer/TidyX/tree/master/TidyTuesday_Explained/035-Rectangles), [Ellis Hughes](https://twitter.com/Ellis_hughes) and [Patrick Ward](https://twitter.com/OSPpatrick)!

A video going through this data can be found on [YouTube](https://bit.ly/TidyX_Ep35). Their original scraping code can be found on [GitHub](https://github.com/thebioengineer/TidyX/blob/master/TidyTuesday_Explained/035-Rectangles/Washington-hikes.R). The Washington Trails Assocation website includes [this leaflet-style map visual](https://www.wta.org/go-outside/map)

Let's setup with the `tidyverse`, a theming package `hrbrthemes`, some web scraping tools, and table building tools.

```{r setup }
suppressPackageStartupMessages({
library(tidyverse)
library(hrbrthemes)
extrafont::loadfonts(quiet = TRUE)
library(rvest)
library(gt)
library(fontawesome)
library(htmltools)

library(ggridges)

})

source(here::here("code","_common.R"),
       verbose = FALSE,
       local = knitr::knit_global())
ggplot2::theme_set(theme_jim(base_size = 12))
```

The original dataset provided by R4DS:

```{r Load}

tt <- tidytuesdayR::tt_load("2020-11-24")

```

```{r Readme, eval = interactive(), include=FALSE}

tt

```

After visiting the Washington Trails Association web site, I had expected to find 3869 hikes. It turns out that the R4DS scraping script skips all trails that lack elevation, gain, length, or rating figures. For more advanced visuals, I decided to go back and re-visit the scraping algorithm to also gather the geospatial coordinates in the sub pages.  The update is here:

```{r scraping}
hike_data_rds <- here::here("data/hike_data.rds")

if (!file.exists(hike_data_rds)) {
  
  hike_pages <- 1:129
  
  hike_data <- lapply( hike_pages - 1 , function(page){
    
    start_int <- 30 * page
    
    page_url <- paste0(
      "https://www.wta.org/go-outside/hikes?b_start:int=",
      start_int
    )
    
    page_html <- read_html(page_url)
    
    page_html %>% 
      
      html_nodes(".search-result-item") %>% 
      
      map(
        function(hike){
          
          hike_name <- hike %>%
            html_nodes(".item-header") %>% 
            html_nodes(".listitem-title") %>% 
            html_nodes("span") %>%  
            html_text()

          hike_page <- hike %>% 
            html_nodes(".item-header") %>% 
            html_nodes(".listitem-title") %>% 
            html_attr("href") 
          
          hike_location <- hike %>% 
            html_nodes(".item-header") %>% 
            html_nodes(".region") %>% 
            html_text()

          hike_desc <- hike %>% 
            html_nodes(".hike-detail") %>% 
            html_nodes(".listing-summary") %>% 
            html_text() %>% 
            stringr::str_trim(side = "both")
          
          hike_features <- hike %>% 
            html_nodes(".hike-detail") %>% 
            html_nodes(".trip-features") %>% 
            html_nodes("img") %>% 
            html_attr("title") %>% 
            list()
          
          tibble(
            name = hike_name,
            location = hike_location,
            features = hike_features,
            description = hike_desc,
            hike_page = hike_page
          )
        }) %>% 
      bind_rows() %>% 

      mutate(stats = map(hike_page, function(hike_page){
          print(hike_page)

          stats_html <- read_html(hike_page) 
          
          length <- stats_html %>% 
                    html_nodes(".hike-stat") %>%
                    html_nodes("#distance") %>%
                    html_nodes("span") %>%
                    html_text()
          
          elevation <- stats_html %>% 
                       html_nodes(".hike-stat") %>% 
                       `[[`(3)
          
          elevation_state <- case_when(
                sum(html_text(elevation) %>% str_detect(c("Gain","Point"))) == 2 ~ "both",
                html_text(elevation) %>% str_detect("Gain") ~ "gain",
                html_text(elevation) %>% str_detect("Point") ~ "highpoint",
                TRUE ~ "neither"
          )
          
          if ( elevation_state == "both") { 
            gain <- elevation %>%
                            html_nodes("span") %>% 
                            html_text() %>% 
                            `[[`(1)
            
            highpoint <- elevation %>%
                            html_nodes("span") %>% 
                            html_text() %>% 
                            `[[`(2)
            
            } else if ( elevation_state == "highpoint") {
             
             highpoint <- elevation %>%
                            html_nodes("span") %>% 
                            html_text() 
             
             gain <- NA_character_
             
           } else if ( elevation_state == "gain") {
             
             gain <- elevation %>%
                            html_nodes("span") %>% 
                            html_text() 
             
             highpoint <- NA_character_
             
           } else if (elevation_state == "neither") {
             highpoint <- NA_character_
             gain <- NA_character_
           }
         
      rating <- stats_html %>% 
                      html_nodes(".hike-stat") %>% 
                      `[[`(4) %>%
                      html_nodes(".current-rating") %>%
                      html_text()
            
      longitude <- if (is_empty(stats_html %>% 
                       html_nodes(".latlong") %>%
                       html_node("span") %>% 
                       html_text)) {NA_character_} else {
        
                      stats_html %>% 
                      html_nodes(".latlong") %>%
                      html_nodes("span") %>% 
                      html_text() %>% 
                      `[[`(1) }
      
      latitude <- if (is_empty(stats_html %>% 
                       html_nodes(".latlong") %>%
                       html_node("span") %>% 
                       html_text)) {NA_character_} else {
        
                      stats_html %>% 
                      html_nodes(".latlong") %>%
                      html_nodes("span") %>% 
                      html_text() %>% 
                      `[[`(2) }
      
    stats <- tibble(
        length = length,
        gain = gain,
        highpoint = highpoint,
        rating = rating,
        longitude = longitude,
        latitude = latitude)
      
    return(stats)
      }))
    
  }) %>% 
    bind_rows() %>% 
    unnest_wider(stats) 
  
hike_data <- hike_data %>% 
    mutate(
      trip = case_when(
      grepl("roundtrip",length) ~ "roundtrip",
      grepl("one-way",length) ~ "one-way",
      grepl("of trails",length) ~ "trails"),
     length_total = as.numeric(gsub("(\\d+[.]\\d+).*","\\1", length)) * ((trip == "one-way") + 1)) %>% 
   mutate(
    across(c(length, gain, highpoint, longitude, latitude, rating), parse_number))  
  write_rds(hike_data, file = hike_data_rds)
  
}else{
  
  hike_data <- read_rds(hike_data_rds)
  
}

```

We will further separate the location field into region and subregion, by the hyphen.

```{r}
hike_data <- hike_data %>% 
   separate(location, into = c("region","subregion"), sep = " -- ", fill = "right") %>% 
   replace_na(list(subregion = "Other")) %>% 
   mutate(across(c(name, region, subregion), as_factor))

```

First, a table of summarized trail data by Region

```{r}
rating_stars <- function(rating, max_rating = 5){
  rounded_rating <- floor(rating + 0.5)
  stars <- lapply(seq_len(max_rating), function(i){
    if (i <= rounded_rating) fontawesome::fa("star", fill = "orange") 
    else fontawesome::fa("star", fill = "grey")
  })
  label <- sprintf("%s out of %s", rating, max_rating)
  div_out <- div(title = label, "aria-label" = label, role = "img", stars)
  
  as.character(div_out) %>% 
    gt::html()
}

hike_data %>% 
  group_by(region, subregion) %>% 
  summarize(total_length = round(sum(length, na.rm = TRUE)),
            average_gain = round(mean(gain, na.rm = TRUE)),
            min_rating = round(min(rating, na.rm = TRUE),1),
            max_rating = round(max(rating, na.rm = TRUE),1),
            number = n(),
            .groups = "drop") %>% 
  mutate(subregion = fct_reorder(subregion, desc(number)),
         min_rating = map(min_rating, rating_stars),
         max_rating = map(max_rating, rating_stars)) %>%    
  select(region, subregion, number, total_length, average_gain, min_rating, max_rating) %>% 
  arrange(region, subregion) %>% 
  group_by(region) %>% 
  gt() %>% 
  tab_header(title = gt::html("<span style='color:darkgreen'>Washington State Hiking Trail Summary by Region</span>")) %>% 
  cols_label(subregion = "Region",
             number = "Number of Trails",
              total_length = md("Total Milage All Trails"),
              average_gain = md("Average Elevation Gain (Feet)"),
              min_rating = md("**Min Rating**"),
              max_rating = md("**Max Rating**")) %>% 
  opt_row_striping() %>% 
  tab_options(data_row.padding = px(3))  
```

Another table, adapted from [Kaustav Sen's post](https://github.com/kaustavSen/tidytuesday/blob/master/2020/week_48_washington_trails.Rmd)

```{r}
thumbs_up <- function(value) {
  icon <- fa("thumbs-up", fill = "#38A605")
  div(icon) %>% 
    as.character() %>% 
    html()
}

bar_chart <- function(value, color = "#DFB37D"){
    glue::glue(
      "<span style=\"display: inline-block; 
      direction: ltr; border-radius: 4px; 
      padding-right: 2px; background-color: {color}; 
      color: {color}; width: {value}%\"> &nbsp; </span>"
    ) %>% 
    as.character() %>% 
    gt::html()
}


hike_data %>% 
  group_by(region) %>% 
  arrange(desc(rating)) %>% 
  slice(1:4) %>% 
  ungroup() %>% 
  select(-description, -gain, -highpoint, -latitude, -longitude,
         -hike_page, -subregion, -length_total, -trip) %>%
  mutate(rating = if_else(is.na(rating),0,rating), 
    length_bar = map(length, bar_chart),
    rating = map(rating, rating_stars),
    feature_present = 1,
    feature_present = map(feature_present, thumbs_up)
  ) %>% 
  unnest_longer(features) %>% 
  mutate(features = str_replace(features, pattern = "/.+", replacement = "")) %>% 
  pivot_wider(
    names_from = features,
    values_from = feature_present
  ) %>%
  group_by(region) %>% 
  arrange(-length) %>% 
  select(name, length, length_bar, rating, everything()) %>% 
  gt() %>% 
  cols_width(
    "name" ~ px(250),
    "length" ~ px(75),
    "length_bar" ~ px(140),
    "rating" ~ px(140),
    4:last_col() ~ px(120)
  ) %>% 
  tab_spanner(
    label = "Features",
    columns = 5:last_col()
  ) %>% 
  cols_label(
    name = "",
    length = md("Length<br/>(in miles)"),
    length_bar = "",
    rating = "Rating"
  ) %>% 
  tab_header(
    title = "Top Hiking Trails to visit in Washington",
    subtitle = "The 4 top rated trails in each of the 11 major regions of Washington State"
  ) %>% 
  tab_source_note(
    source_note = md("**Data:** Washington Trails Association | **Table:** Jim Gruman")
  ) %>% 
  tab_style(
    style = cell_text(
      font = google_font("Alegreya Sans SC"),
      weight = "bold"
    ),
    locations = list(cells_row_groups(), cells_column_spanners("Features"))
  ) %>% 
  tab_style(
    style = cell_fill(
      color = "#DFB37D",
      alpha = 0.5
    ),
    locations = cells_row_groups()
  ) %>% 
  tab_style(
    style = cell_text(
      font = google_font("Alegreya Sans"),
      v_align = "bottom",
      align = "center"
    ),
    locations = cells_column_labels(columns = everything())
  ) %>% 
  tab_style(
    style = cell_text(
      font = google_font("Alegreya Sans"),
    ),
    locations = cells_body("name")
  ) %>% 
  tab_style(
    style = cell_text(
      font = google_font("Alegreya Sans"),
      align = "left",
      size = px(30)
    ),
    locations = cells_title("subtitle")
  ) %>% 
  tab_style(
    style = cell_text(
      size = "medium"
    ),
    locations = cells_body(columns = 4:last_col())
  ) %>%
  tab_style(
    style = cell_text(
      align = "left"
    ),
    locations = cells_body(columns = "length_bar")
  ) %>%
  tab_style(
    style = cell_text(
      font = google_font("Alegreya Sans SC"),
      align = "left",
      weight = "bold",
      size = px(50)
    ),
    locations = cells_title("title")
  ) %>% 
  opt_table_font(font = google_font("Fira Mono")) %>% 
  tab_options(
    table.border.top.color = "white",
  ) 
```

What is the distribution of hiking trail lengths around the state regions?

```{r}

hike_data %>% 
  filter(!is.na(length)) %>% 
  ggplot(aes(length, region, fill = region)) +
  geom_density_ridges(alpha = 0.9, show.legend = FALSE) +
  scale_x_log10(labels = scales::comma_format(accuracy = 1)) +
  coord_cartesian(clip = "off") +
  labs(x = "Trail Lengths in Miles, Log Scale", y = NULL, 
       title = "Trail Lenths Available by Region", 
       subtitle = "The Cascades have more of the Longer Trails", 
       caption = "@jim_gruman | Data: TidyX") 
```

Let's take a closer look at the features of each trail and the relationship with length:

```{r}
hikes_complete <- hike_data %>%
    filter(length > 0.1) %>%
    mutate(hikeID = row_number(),
           length = log10(length),
           length = length %/% 0.2 * 0.2) %>%
    select(hikeID, length, features) %>%
    unnest(features) %>%
    count(length, features) %>%
    complete(length, features, fill = list(n = 0)) %>%
    group_by(length) %>%
    mutate(length_total = sum(n)) %>%
    ungroup()

hikes_complete
```

```{r}
slopes <- hikes_complete %>%
    nest_by(features) %>%
    mutate(model = list(glm(cbind(n, length_total) ~ length, 
                            family = "binomial", 
                            data = data))) %>%
    summarize(broom::tidy(model)) %>%
    ungroup %>%
    filter(term == "length") %>%
    mutate(p.value = p.adjust(p.value)) %>%
    filter(p.value < 0.05) %>%
    arrange(estimate)
```


```{r}
hikes_complete %>%
    mutate(length_percent = n / length_total) %>%
    inner_join(slopes) %>%
    mutate(features = fct_reorder(features, -estimate)) %>%
    ggplot(aes(length, length_percent)) +
    geom_smooth(method = "lm") +
    geom_point() +
    facet_wrap(~features) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Length of each trail (log10 scale)",
         y = "% of trails with each feature",
         title = "Which features of Washington hikes are associated with longer trails?",
         caption = "@jim_gruman | Data: TidyX")
```

